\subsection{The Role of the Universal Bound \texorpdfstring{$c_\chi$}{cχ}}
  \label{subsec:role-of-cchi}

  A central structural element of the Cosmochrony framework is the existence of a
  universal invariant bound, denoted $c_\chi$, defined at the level of the pre-temporal
  $\chi$ substrate.
  This bound does not correspond to a signal velocity, nor to the propagation of any
  field, excitation, or information in spacetime.
  Instead, $c_\chi$ characterizes an absolute structural limit on the degree to which
  relational information can be locally confined within admissible configurations of
  the $\chi$ substrate.

  At the fundamental level, $c_\chi$ is non-metric and non-temporal.
  It is not associated with distances, durations, lightcones, or causal ordering, since
  none of these notions are defined prior to projection.
  Rather, $c_\chi$ expresses a maximal admissible rate of structural ordering: beyond
  this limit, further confinement of relational degrees of freedom becomes impossible
  and relaxation necessarily occurs.

  Crucially, $c_\chi$ is not itself an observable quantity.
  It acquires operational meaning only through projection, when configurations of
  $\chi$ admit a locally injective representation in terms of effective spacetime
  variables.
  In such projectable regimes, the invariant structural bound $c_\chi$ manifests as an
  effective causal constraint $c$, defined through the maximal admissible local ordering
  rate of projected configurations:
  \[
    c \;\equiv\; \Pi(c_\chi),
  \]
  where $\Pi$ denotes the projection from the pre-geometric relational substrate to an
  effective spacetime description.

  The constant $c$, which coincides numerically with the observed speed of light, thus
  has a strictly derivative status.
  It is not an independent postulate, but the spacetime expression of the deeper
  structural bound $c_\chi$ once notions of locality, duration, and causal ordering
  become meaningful.
  All effective causal constraints appearing in projected descriptions are inherited
  from this underlying invariant bound.

  Accordingly, the local constraint on the effective relaxation functional,
  \[
    \left| \mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}} \right| \le c ,
  \]
  should be understood as the projected manifestation of the more fundamental structural
  limitation imposed by $c_\chi$.
  No causal or dynamical principle is imposed directly at the level of spacetime;
  effective causality arises solely as a consequence of the bounded projective
  realization of relational ordering.

  This distinction becomes essential in strong-gravity or near-deprojection regimes.
  While the effective constant $c$ may lose its geometric or causal interpretation when
  projection ceases to be locally injective, the structural bound $c_\chi$ remains
  invariant.
  The breakdown of spacetime description therefore signals not a violation of
  causality, but the loss of representability of relational configurations within a
  spacetime framework.

  In summary, $c_\chi$ defines a universal, pre-temporal structural bound governing the
  admissibility of $\chi$ configurations, while $c$ represents its effective spacetime
  manifestation.
  The latter inherits its numerical value and universality entirely from the former,
  ensuring conceptual continuity between the pre-geometric substrate and emergent
  relativistic causality.

  \paragraph{Status of Universal Constants.}

    Within the Cosmochrony framework, quantities conventionally treated as fundamental
    constants of nature acquire a derivative and structural status.
    In particular, the constants $c$ and $\hbar$ are not introduced as independent
    postulates, but arise as \textbf{emergent invariants} associated with the projection
    from the pre-geometric $\chi$ substrate to effective observables.

    The effective speed $c$ reflects the maximal admissible rate of relational ordering
    once a locally injective spacetime description becomes available, as discussed above.
    By contrast, $\hbar$ characterizes the minimal admissible unit of reprojection or
    spectral action in the effective description, encoding the irreducible granularity
    with which relational configurations can be operationally resolved
    (see Section~\ref{subsec:A12-potential-role_v171}).

    From this perspective, relativistic causality and quantum discreteness do not stem
    from independent foundational principles, but represent complementary manifestations
    of the same underlying projective constraints acting on the $\chi$ substrate.

    The invariant bound $c_\chi$ thus constrains the maximal rate at which relational
    configurations can be ordered once projected.
    A distinct but complementary invariant governs the minimal granularity with which
    such configurations can be reprojected into effective observables.
    This second structural bound is the subject of the next section.
