\begin{abstract}
  Cosmochrony is a pre-geometric relational framework in which time,
  spacetime geometry, and physical observables emerge from an ordered
  sequence of effective reprojections constrained by a single underlying
  substrate, denoted~$\chi$.
  The substrate~$\chi$ is static and encodes only admissibility
  constraints; it does not carry dynamical evolution.
  Observable dynamics corresponds to the iterated reprojection
  \begin{equation*}
    U_{t+1} = \Pi \circ \sigma(U_t),
  \end{equation*}
  where $\Pi$ is a fixed, generically non-injective projection,
  $\sigma : O \to \Omega$ is an admissible section selecting a
  configuration $\omega_t = \sigma(U_t)$ within the fiber
  $\Pi^{-1}(U_t)$ under the constraints of~$\chi$, and the
  composition $\Pi \circ \sigma$ acts as a deforming mirror:
  $\sigma$ lifts the current state back into the configuration space,
  and $\Pi$ reflects it into the next observable state.
  Because $\sigma$ is not the inverse of~$\Pi$, the reflection is
  deformed by two distinct mechanisms.
  First, the admissibility constraints encoded in~$\chi$ govern which
  configuration $\omega_t$ within the fiber $\Pi^{-1}(U_t)$ is selected:
  the selection is structural, not random.
  Second, the projection noise: distinct configurations within the same
  fiber never overlap perfectly under~$\Pi$; their residual differences,
  inaccessible to the effective description, appear as fluctuations in
  $U_{t+1}$.
  It is this irreducible residue that guarantees $U_{t+1} \neq U_t$ and
  makes the sequence irreversible: a state with no residue would imply
  configurations projecting identically with no relational difference,
  leaving no motive for a next state---the structural analogue of absolute
  zero.
  The substrate~$\chi$ constrains the admissibility of~$\sigma$ but
  does not supply dynamical content: it governs what can be reflected,
  not what is.

  The framework adopts a principle of ontological minimalism,
  postulating no fundamental spacetime, metric, or quantum structure.
  Effective geometric and dynamical descriptions arise solely through
  coarse-grained projections of configurations admissible under~$\chi$.

  The admissible sector is characterized by a spectral structure
  defined by a relational operator on the configuration space of~$\chi$.
  Physical observables correspond to projectable modes selected by
  a spectral admissibility condition.
  Temporal order is identified with the monotonic ordering of the
  reprojection sequence, not with an evolution of the substrate itself.
  Spatial relations and curvature emerge as effective summaries of
  correlations within projectable configurations.

  Bounded-response regimes arise from saturation of admissibility
  constraints, yielding Born--Infeld-like effective dynamics without
  introducing additional fundamental fields.
  In this context, the speed of light and Planck's constant appear as
  complementary bounds on projectability and resolvability rather
  than as primitive constants.

  Matter, inertia, and electric charge correspond to stable or
  metastable invariants of the effective reprojection sequence,
  associated with localized spectral configurations.
  The topology of the projection fiber further constrains admissible
  configurations: electric charge is identified with a winding invariant
  of the $U(1)$ fiber, isolated magnetic monopoles are excluded in the
  canonical admissible sector, and parity asymmetry arises from
  projective chirality induced by non-injectivity.

  Quantum indeterminacy, entanglement, and violations of Bell
  inequalities arise as structural consequences of the non-injective
  character of~$\Pi$: the fiber $\Pi^{-1}(U_t)$ generically contains
  multiple admissible configurations whose joint projection is
  non-factorizable, without requiring superluminal influences or
  fundamental wavefunction collapse.

  At macroscopic scales, gravitation emerges as a collective effect of
  constrained reprojection, recovering general-relativistic behavior in
  appropriate regimes.
  At galactic and cosmological scales, the same bounded-response
  mechanism accounts for flat rotation curves and
  environment-dependent Hubble parameter inference without invoking
  dark matter particles or a fundamental dark energy component.

  Cosmochrony is proposed as a structural unification framework,
  clarifying how time, geometry, quantum phenomena, and cosmological
  behavior arise from a common pre-geometric substrate governed solely
  by admissibility constraints on an iterated non-injective projection.
\end{abstract}
