\subsection{Ontological Minimalism and the Definition of Fundamentality}
  \label{subsec:minimalism-as-a-guiding-principle}

  The minimalism adopted in this work is not merely a methodological or aesthetic
  preference but a constitutive requirement.
  To be fundamental is to depend on no prior physical structure: no background
  spacetime, no independently postulated state space, no external dynamics, and no
  collection of primitive fields whose own origin would remain unexplained.
  Minimising the assumed origin of physical structure is therefore not only a matter
  of parsimony but an operational criterion of fundamentality.
  Every independent postulate leaves a residual explanatory debt, whereas a
  foundational description should introduce only the structure required for physical
  distinctions and their possible continuation.
  Cosmochrony accordingly takes as its primitive the local structure of admissible
  transitions between observable states.
  Given an observable state~$O_{n-1}$, the primitive question is which successor
  directions are admissible from it.
  This structure is governed by local projective admissibility, structural
  non-injectivity, non-premature selection, and projection locking.
  It does not presuppose a background manifold, a Hilbert space, a fundamental
  temporal parameter, or an independently evolving substrate.
  Non-injectivity is not appended to this primitive as an optional modelling choice.
  Any genuinely emergent description must identify distinctions present at one level
  that are no longer independently resolvable at the effective level.
  An injective and structurally faithful correspondence would preserve the complete
  structure of the preceding level and would therefore amount only to a relabelling,
  not to emergence.
  The projective relation is consequently generically non-injective wherever the
  observable description is genuinely distinct from the structure from which it
  emerges.
  The notation~$\Pi_n$ compactly represents this local structure of admissible
  identification.
  It is not an additional physical entity or a dynamical operation acting on a
  pre-existing substrate.
  Likewise, when the global symbol~$\chi$ is used as a representation of the
  pre-geometric relational structure, it is static and carries no independently
  postulated evolution.
  Observable evolution is instead the induced ordering of admissible transitions.
  Time, space, quantum structure, gauge structure, matter, and gravitation are
  therefore not independent primitives.
  They must arise as effective invariants or consequences of the admissibility
  structure.
  Temporal ordering is induced by the oriented succession of admissible projections;
  spatial and geometric relations arise from the stable structure preserved under
  projection; and quantum and gauge descriptions arise from the unresolved,
  non-factorisable fibre structure associated with non-injective admissibility.
