\subsection{Key Conceptual Terms}
  \label{subsec:key-conceptual-terms}

  \paragraph{Admissibility envelope.}
    Under the bounded-flux constraint $|\partial_t \chi| \le c_\chi$, a Laplacian eigenmode
    with eigenvalue $\lambda_n$ admits a maximal effective amplitude
    \[
      A_n^{\max} = \frac{c_\chi}{\sqrt{\lambda_n}}.
    \]
    This relation defines the \emph{admissibility envelope} governing the maximal
    projectable amplitude of each spectral sector.

  \paragraph{Admissibility ratio.}
    The \emph{admissibility ratio} compares the maximal admissible amplitudes of two
    spectral sectors,
    \[
      R_{p/q} = \frac{A_p^{\max}}{A_q^{\max}}
      = \sqrt{\frac{\lambda_q}{\lambda_p}} .
    \]
    This ratio quantifies the relative dynamical preference between spectral sectors
    under the bounded-flux constraint.
    In quaternionic Cayley graphs such as that of $Q_8$, the ratio
    $\sqrt{4/3}$ between the non-abelian and abelian sectors provides the minimal
    instance of this hierarchy.

  \paragraph{Character-vanishing condition.}
    The condition
    \[
      \chi_\rho(s) = 0 \quad \forall s \in S
    \]
    is referred to as the \emph{character-vanishing condition}.
    It identifies generator sets whose associated $SU(2)$ representation consists entirely
    of traceless elements.
    This condition plays a central role in the classification of admissible spectral sectors,
    since it ensures that the adjacency contribution $\mu_\rho$ vanishes and the Laplacian
    eigenvalue satisfies $\lambda_\rho = |S|$.

  \paragraph{Cumulative projected capacity \texorpdfstring{$I(n)$}{I(n)}.}
    For a conjugate pair $(c,\,q-c)$ of Weil representations on
    $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$, the cumulative projected capacity is
    \[
      I(n) \;=\; \sigma_{\mathrm{pair}}(0) - \sigma_{\mathrm{pair}}(n),
    \]
    where $n$ indexes BFS shell depth and $\sigma_{\mathrm{pair}}(n)$ is the
    residual projective capacity at depth~$n$.
    $I(n)$ measures the spectral structure already stably projected up to depth~$n$.
    Its strict non-decreasing character along tested BFS trajectories (Proposition~1
    of~\cite{Beau2026a28}) provides the canonical numerical realization of the
    projective temporal ordering~$\tau$: a larger $I(n)$ corresponds to a later
    effective state.
    \emph{Status:} numerical result for $q \in \{29,61,101,151\}$; analytical
    derivation of the monotonicity remains an open problem.

  \paragraph{Capacity exponent \texorpdfstring{$\delta_{\mathrm{pair}}$}{δpair}.}
    The power-law exponent governing the decay of the residual projective capacity
    $\sigma_{\mathrm{pair}}(n) \sim n^{-\delta_{\mathrm{pair}}}$ along the BFS
    cascade.
    It is the primary spectral observable of the O-series and connects to the
    cascade exponent $\beta^*$ governing the fermion mass hierarchy via
    the structural relation established in~O24.
    Current numerical values: $\delta_{\mathrm{pair}} \approx 8.89$ ($q=29$),
    $9.98$ ($q=61$), $9.55$ ($q=101$).

  \paragraph{Energy.}
    Energy is an effective quantity measuring the operational cost associated
    with modifying a projected configuration while preserving coherence under
    admissible updates.
    This cost reflects the mobilization of the underlying projective capacity
    of the $\chi$ substrate under projection, but energy itself is not a
    fundamental property of $\chi$.
    Standard conservation laws remain valid at the effective level.

  \paragraph{Fluctuations.}
    Local stochastic modulations of $\chi$ configurations that affect event timing
    and localization without altering underlying topological constraints.

  \paragraph{Gram rigidity.}
    Let $\{\vec u_s\}_{s \in S}$ denote the unit axes associated with neutral generators.
    The pairwise geometry of these axes is encoded by the Gram matrix
    \[
      G_{st} = \vec u_s \cdot \vec u_t .
    \]
    Through the character identity
    \[
      G_{st} = -\frac{1}{2}\chi_\rho(st),
    \]
    the entire geometric configuration is determined by representation-theoretic data.
    \emph{Gram rigidity} refers to the fact that admissible neutral configurations are
    severely constrained by this identity, leading to a discrete classification of
    compatible finite subgroups of $SU(2)$.

  \paragraph{Matter.}
    Stable topological configurations of $\chi$ whose persistence gives rise to
    particle-like behavior and inertial properties.

  \paragraph{Measurement.}
    A localized interaction that selects a specific effective realization
    within a non-injective projection.
    Measurement does not induce a fundamental collapse, but resolves an
    equivalence class of projected configurations into a stable operational
    outcome.

  \paragraph{Neutral generator.}
    A generator $s \in S$ of a finite group representation $\rho : G \to SU(2)$ is said to be
    \emph{neutral} if its character vanishes,
    \[
      \chi_\rho(s) = 0.
    \]
    Neutral generators correspond to traceless elements of $SU(2)$ and therefore to
    half-turn rotations in the associated $SO(3)$ representation.
    Within the spectral admissibility programme, neutrality ensures that the corresponding
    sector does not receive constructive alignment from the generating set and therefore
    admits the maximal admissibility window compatible with bounded relational flux.

  \paragraph{One-way activation.}
    The property that a spectral mode, once stably projected at BFS depth~$n$,
    does not leave the Gram--Schmidt span at any subsequent depth $n' > n$.
    Formally, $\delta r_n \geq 0$ at every step, so that $I(n)$ is non-decreasing.
    One-way activation is the spectral mechanism underlying the irre\-versibility
    of the projective temporal ordering: it provides a structural arrow of time
    without invoking thermodynamic entropy or a background temporal coordinate.
    \emph{Status:} established numerically for $q \in \{29,61,101,151\}$;
    the analytical proof from BFS admissibility constraints remains open.

  \paragraph{Projective temporal ordering.}
    The ordering relation on effective states defined by
    \[
      U_t \prec U_{t+1} \;\Longleftrightarrow\; I(U_t) \leq I(U_{t+1}).
    \]
    This ordering is not a dynamics on $\chi$ but a relational structure induced
    by the non-injective projection~$\Pi$: it reorders admissible fibre selections
    within the fixed observable rank, without introducing additional observable
    degrees of freedom.
    It is compatible with and complementary to the observable rank stability
    established in~O24.
    See~\cite{Beau2026a28} for the numerical evidence and formal statement.

  \paragraph{Reprojection sequence.}
    The iterated map
    \[
      U_{t+1} = \Pi \circ \sigma(U_t),
    \]
    where $\sigma(U_t) \in \Pi^{-1}(U_t)$ selects an admissible configuration
    within the fibre compatible with the current effective state~$U_t$.
    The reprojection sequence is the formal carrier of effective dynamics in
    Cosmochrony: all observable evolution arises from successive admissible
    fibre selections, not from any evolution of the static substrate~$\chi$.
    Successive steps correspond to distinct admissible realisations drawn from
    different regions of the fibres; the effective time parameter $\tau$ indexes
    this sequence via the monotone accumulation of $I(n)$.

  \paragraph{Residual projective capacity \texorpdfstring{$\sigma_{\mathrm{pair}}(n)$}{σpair(n)}.}
    For a conjugate pair $(c,\,q-c)$, the residual projective capacity at BFS
    depth~$n$ is
    \[
      \sigma_{\mathrm{pair}}(n)
      = \sigma_c(n)\cdot\sigma_{q-c}(n),
      \quad
      \sigma_c(n) = \frac{\delta r_n}{|S_n|},
    \]
    where $\delta r_n$ counts new linearly independent Weil fingerprint vectors
    in BFS shell~$S_n$ and $|S_n|$ is the shell size.
    $\sigma_{\mathrm{pair}}(n)$ decays monotonically toward zero as the spectral
    capacity of the pair is exhausted; its complement is the cumulative projected
    capacity $I(n)$.
    The Born--Infeld saturation plateau is reached when $\sigma_{\mathrm{pair}}(n)
\approx 0$.

  \paragraph{Spectral capacity.}
    For a finite group $G$ with symmetric generating set $S$, the
    \emph{spectral capacity} functional is defined as
    \[
      C(G,S) =
      \sum_{\rho \in \widehat G,\; \lambda_\rho > 0}
      \frac{(\dim \rho)^2}{\sqrt{\lambda_\rho}},
    \]
    where $\lambda_\rho$ are the Laplacian eigenvalues associated with the irreducible
    representation sectors.
    The functional aggregates the admissible amplitude windows of all spectral sectors
    weighted by their Peter--Weyl multiplicities.
    Spectral capacity provides a global measure of how efficiently a relational structure
    supports dynamically admissible modes under bounded projective flux.

  \paragraph{Probability.}
    An emergent descriptor reflecting the multiplicity of projected realizations
    compatible with a given relational configuration and effective state.
    Probability is projective rather than fundamental, and does not correspond
    to intrinsic randomness of the substrate.

  \paragraph{Projective capacity.}
    The available ``room'' within the admissible projective structure for
    further structural reorganization.
    At the spectral level it is quantified by $\sigma_{\mathrm{pair}}(n)$
    and its integral $I(n)$ along the BFS cascade.
    At the effective level it governs the coupling $g(x,t) = \gamma/\Sigma(x,t)$
    in the nonlinear Schrödinger correction (§B.14) and the differential between
    constrained and unconstrained vacuum configurations (Casimir effect).
    Projective capacity is contextual and non-extensive; it replaces the
    notion of ``relaxation capacity of $\chi$'' used in earlier formulations
    of the framework.

  \paragraph{Spacetime.}
    An emergent relational structure arising from large-scale configurations of
    the $\chi$ substrate.
    Its metric description remains valid within its domain of applicability.

  \paragraph{Time.}
    An effective ordering parameter associated with the local rate of projective
    ordering.
    Operational and relativistic notions of time are recovered without modification.
    The canonical spectral realization is the cumulative projected capacity
    $I(n)$ (see above).

  \paragraph{Projective transmittance (gauge interpretation).}
    An effective, context-dependent measure of how efficiently admissible projective
    flux is transmitted through a given projected configuration.
    In Cosmochrony, gauge structure can be interpreted as a parametrization of
    transmittance variations across the projection fiber, rather than as a fundamental
    interaction field.

  \paragraph{Wavefunction.}
    An effective statistical representation of the dynamics and topology of the
    $\chi$ substrate.
    It has no fundamental ontological status.

  \paragraph{Wave--Particle Duality.}
    A manifestation of interaction-induced changes in the local configuration of
    $\chi$, producing localized particle-like behavior from an underlying
    wave-like substrate.

  \paragraph[$w$]{Winding number (fiber topological class).}
    $w$ labels the winding class of a configuration in the projected $U(1)$ fiber sector.
    A $w=1$ configuration corresponds to an elementary charged excitation in the effective description.
    A $w=2$ configuration denotes a composite topological class supporting a collective phase degree of freedom.
    The winding number is a topological invariant of the projected description and does not constitute a new dynamical
    postulate.
    It is used to classify admissible coherent regimes and their stability properties.
