% ----------------------------------------------------------------------------
% Section 2.5 --- Structural Principles and Projective Regimes
% From former ยง3.5
% ----------------------------------------------------------------------------
\subsection{Structural Principles and Projective Regimes}
\label{subsec:structural-principles-and-projective-regimes}

All effective physical observables arise through a relational
projection~$\Pi$ from the underlying configuration space~$\Omega$ of
the~$\chi$ substrate.
Effective descriptions are constrained by three structural principles that
govern the admissibility of projected regimes.

\paragraph{Principle I: Substratic Locality and Bounded Relaxation.}
The fundamental relaxation dynamics of~$\chi$ is strictly local and subject
to universal bounds.
The transport of relational relaxation admits a maximal admissible flux,
constraining the rate at which structural information can be redistributed.
This bound arises from the intrinsic stability conditions of~$\chi$ itself.

\paragraph{Principle II: Non-Injective Projective Realization.}
The projection $\Pi : \Omega \rightarrow O$ from substratic configurations
to effective observables is generically non-injective.
Distinct configurations of~$\chi$ may be structurally identified at the
level of observable descriptions.
This implies that effective descriptions need not admit a factorisable or
locally complete representation, even when the underlying dynamics remains
strictly local.

\paragraph{Principle III: Projective Compensation.}
Whenever $\Pi$ fails to resolve the full relational complexity of~$\chi$,
effective descriptions compensate through the inflation of effective
parameters---temperature, curvature, horizon structure---encoding
unresolved relational structure within a reduced descriptive framework.
These quantities function as Lagrange multipliers, not as additional
fundamental degrees of freedom.

Together, these principles delineate distinct projective regimes.
When the projection is approximately injective and relaxation is far from
saturation, standard local and geometric descriptions apply.
Near structural saturation or strong non-injectivity, effective parameters
grow large, signaling the breakdown of spacetime-based descriptions.
