\subsection{Conceptual Implications and Open Challenges}
  \label{subsec:conceptual-implications-and-open-challenges}

  Temporal ordering arises from the monotonic relaxation of~$\chi$; energy quantifies residual capacity to resist this
  relaxation; irreversibility reflects its progressive exhaustion.
  Time, energy, and irreversibility are not independent axioms but complementary effective descriptions of the same
  relational dynamics.

  Open challenges include the quantitative reconstruction of CMB anisotropies from early-time $\chi$ dynamics, the detailed
  treatment of non-equilibrium decoherence and reprojection, the emergence of gauge symmetries from topological
  features, the explicit derivation of the full flavor hierarchy and the quantitative determination of the CP-violating
  invariant from first-principles $\chi$ dynamics, and the long-term stability of solitonic configurations under extreme
  conditions.
  Progress will require large-scale simulations of $\chi$ dynamics, discretized lattice realizations, and targeted
  experimental tests.

  \paragraph{Possible Cosmological Implications of Flavor Misalignment.}

    Within the present framework, the Jarlskog invariant $\mathcal{J}$
    is interpreted as a structural residue of quasi-commutative
    sectorial restrictions of the projection fiber~$\Pi$
    (Appendix~\ref{subsec:cp-asymmetry-and-chiral-selection}).
    In the effective low-energy regime, this invariant is treated as
    a stable characteristic of the stabilized flavor manifold.

    Because the operators $\Pi_u$ and $\Pi_d$ are defined as
    restrictions of the global non-injective mapping,
    their algebraic structure is ultimately rooted in the global
    state of the relational substrate~$\chi$.
    In extreme regimes approaching global saturation---such as the
    early Universe---the admissible spectral subspace of the fiber
    could differ from its present quasi-stationary configuration.

    Whether such regimes might modify the effective structure of
    flavor misalignment remains an open question.
    No quantitative model is presently derived.
    Establishing a predictive relation between $\mathcal{J}$ and
    cosmological relaxation conditions constitutes a significant
    theoretical challenge for future work.
