\subsection{Initial Conditions and Global Structure}
  \label{subsec:initial-conditions-and-global-structure}

  The Cosmochrony framework does not postulate initial conditions in the conventional
  temporal sense.
  Instead, it assumes that the relational substrate $\chi$ admits a minimal admissible
  ordering state, denoted $\chi_0$, corresponding to configurations of maximal effective
  relaxation density.
  This reference state does not represent an initial condition imposed at a distinguished
  moment in time, but a structural boundary of admissible projected descriptions,
  characterizing the earliest configurations that can meaningfully admit an effective
  ordering interpretation.

  In effective geometric regimes, the characteristic scale associated with projected
  descriptions in the vicinity of $\chi_0$ coincides numerically with the Planck scale.
  This correspondence reflects the loss of projectability and the breakdown of
  coarse-grained spacetime descriptions below this regime, rather than the presence of a
  fundamental cutoff, microscopic discreteness, or underlying spacetime lattice.

  From this perspective, cosmic history is interpreted as the progressive and irreversible
  ordering of projected $\chi$ configurations away from this minimal admissible state.
  No spacetime singularity is required at the fundamental level.
  Apparent singular behavior arises only when classical notions of time, distance, or
  curvature are extrapolated beyond the regime in which projected $\chi$ configurations
  admit a stable geometric and causal interpretation.

  \paragraph{Ontological poverty and the arrow of complexity.}
    The minimal admissible state $\chi_0$ corresponds to a regime of \emph{ontological
    poverty}: only a small class of simple, highly coherent configurations can be
    projected.
    Early configurations were not merely simpler realizations of a fixed blueprint,
    but belonged to a drastically restricted admissible space, incapable of
    supporting the relational richness that emerged only through subsequent relaxation.

    As relaxation proceeds, this admissible space expands, enabling the emergence of
    hierarchical, localized, and complex structures.
    Cosmological time is therefore simultaneously the arrow of entropy and the arrow
    of increasing descriptive richness.
    The complexity observed today was not encoded or prefigured in advance.

    In this sense, Cosmochrony is fundamentally non-teleological:
    the universe does not unfold a prewritten complexity, but acquires descriptive
    richness only as the space of admissible configurations itself expands.
    This expansion can be characterized graph-theoretically in terms of the spectral
    properties of the relational connectivity; the detailed formalism is developed in
    Appendix~\ref{app:lowell_attenuation}.

  The global structure of admissible projected descriptions is therefore constrained by
  the ordering properties of the underlying $\chi$ substrate, rather than by arbitrarily
  specified initial data or boundary conditions.
  In the following section, we derive a minimal effective dynamical equation governing
  the ordering of projected $\chi$ configurations and explore its immediate physical
  consequences.
