\clearpage
\section{Definition and Fundamental Properties of the $\chi$ Substrate}
  \label{sec:definition-and-fundamental-properties-of-the-chi-field}

  Having outlined the ontological and conceptual principles underlying Cosmochrony, we now
  introduce the single fundamental entity at the core of the framework.
  This section is devoted to defining the pre-geometric relational substrate $\chi$ and to
  clarifying its role as the primitive ontological basis from which effective notions of
  spacetime, dynamics, and physical observables emerge.

  The purpose of this section is not to assume a pre-existing spacetime structure, but to
  identify the minimal structural properties required of $\chi$ in order to recover, in
  appropriate regimes, effective descriptions of time, space, metric geometry, and field
  dynamics.
  Accordingly, $\chi$ is introduced independently of any spacetime coordinates, metric
  structure, or background manifold, and is related to geometric notions only through
  non-injective, coarse-grained projections that become meaningful once a stable spacetime
  interpretation is supported.

  Throughout this section, the use of variational principles, Lagrangian formulations, or
  metric-based expressions does not imply that spacetime or a four-dimensional manifold is
  fundamental.
  Such formalisms are employed strictly as effective descriptive tools, valid only in
  projectable regimes where configurations of $\chi$ admit a quasi-stable geometric and
  causal interpretation.
  They should therefore be understood as post-hoc representations of underlying
  pre-geometric dynamics, not as primary postulates of the theory.

  We begin by providing a unified conceptual definition of the $\chi$ substrate and its
  physical interpretation, emphasizing its ontological distinction from conventional
  fields.
  The subsequent subsections progressively introduce more structured effective
  descriptions—including scalar, Lagrangian, and metric formulations—which become
  applicable only once projected $\chi$ configurations support a coherent spacetime
  regime.

  \input{03-chi-field/01-definition-and-fundamental-properties-of-the-chi-field}
  \input{03-chi-field/02-the-geometric-effective-action-and-lagrangians-of-cosmochrony-Lcc}
  \input{03-chi-field/03-physical-interpretation}
  \input{03-chi-field/04-monotonicity-and-arrow-of-time}
  \input{03-chi-field/05-local-relaxation-speed}
  \input{03-chi-field/06-relation-to-conventional-fields}
  \input{03-chi-field/07-initial-conditions-and-global-structure}
