\subsection{Non-Commutativity as a Source of Mass}
  \label{sec:noncommutativity-mass}

  Torsion acts as a dynamical constraint that competes with relaxation transport
  through the projection fiber, and may obstruct the existence of a common spectral
  description for diffusion and winding-induced constraints.

  \subsubsection*{Inhibition of Relaxation}
    \label{sec:inhibition-relaxation}

    Let $\Omega_w$ be the effective torsion operator associated with winding
    number~$w$. For the fundamental lepton configuration ($w=1$), relaxation and torsion
    remain spectrally compatible:
    \begin{equation}
      [\Delta^{(0)}_G,\Omega_{1}] = 0 .
    \end{equation}
    For higher-winding configurations ($w\ge 2$), torsional constraints become
    spectrally frustrated:
    \begin{equation}
      [\Delta^{(0)}_G,\Omega_{w}] \neq 0
      \qquad (w\ge 2).
    \end{equation}
    The absence of a common eigenbasis prevents uniform relaxation and induces
    anisotropic spectral compression within the admissible subspace, manifesting
    macroscopically as amplified inertial mass.

    A convenient global diagnostic of this distortion is the torsional action
    \begin{equation}
      \mathcal{A}(w) \;\equiv\;
      \frac12 \ln\!\left(
                     \frac{\det_\zeta(\Delta^{(0)}_G+\Omega_{w})}
                     {\det_\zeta(\Delta^{(0)}_G)}
      \right),
      \label{eq:Aw-fredholm}
    \end{equation}
    defined via zeta-regularized determinants. This invariant quantifies global spectral
    deformation, but does not by itself encode dynamical stability selection.

  \subsubsection*{The Pisano Ratio as a Dynamical Stability Criterion}
    \label{sec:pisano-ratio}

    For $w=2$, the first spectral multiplet of $S^3$ admits a natural Hopf-compatible
    decomposition into a fiber-aligned singlet and a transverse triplet, inducing a splitting
    \begin{equation}
      \Pi = \Pi_{\parallel} \oplus \Pi_{\perp}.
      \label{eq:Pi-splitting}
    \end{equation}
    Crucially, the relevant quantities are not the static eigenvalues of
    $\Delta^{(0)}_G+\Omega_2$, but the effective rotational frequencies generated by the
    torsional operator within this reduced multiplet. Denote by $\lambda_\parallel$ and
    $\lambda_\perp$ the two effective frequencies associated with torsion-induced evolution
    in $\Pi_\parallel$ and $\Pi_\perp$.

    Dynamical stability selects the most irrational admissible ratio between these
    frequencies, in analogy with Kolmogorov--Arnold--Moser (KAM) criteria:
    \begin{equation}
      \frac{\lambda_{\parallel}}{\lambda_{\perp}}
      \;=\; \varphi,
      \label{eq:pisano-kam}
    \end{equation}
    where $\varphi=(1+\sqrt5)/2$ is the golden ratio. Equation~\eqref{eq:pisano-kam} is
    therefore a dynamical selection principle imposed on the torsional generator $\Omega_2$
    within the first multiplet, and should not be interpreted as a generic spectral identity
    of $\Delta^{(0)}_G+\Omega_2$ in the weak-torsion regime.

  \subsubsection*{Leptonic Spectrum Synthesis}
    \label{sec:leptonic-synthesis}

    For the muon, non-commutative torsion in the $w=2$ sector combines:
    (i) the topologically fixed spectral ratio $\lambda_2/\lambda_1 = 8/3$,
    (ii) the dynamical Pisano selection $\lambda_\parallel/\lambda_\perp = \varphi$,
    (iii) the universal coupling factor $3/(2\alpha)$,
    yielding the parameter-free prediction
    \begin{equation}
      \boxed{
        \frac{m_\mu}{m_e}
        = \sqrt{\frac{\lambda_{2}}{\lambda_{1}}}
        \cdot \frac{3}{2\alpha}
        \cdot \frac{1}{\varphi}
      }
      \qquad\text{with}\qquad
      \frac{\lambda_2}{\lambda_1} = \frac{8}{3}.
      \label{eq:muon-ratio-final}
    \end{equation}
    The robustness of $\lambda_2/\lambda_1=8/3$ is demonstrated independently in
    Appendix~D.7, while the Pisano factor arises from a dynamical stability condition
    applied to the torsional generator within the first spectral multiplet.

    \paragraph{Transition.}
      Having identified non-commutativity as the universal obstruction for $w\ge 2$, the next
      question is whether higher-winding sectors remain merely anisotropic (as for $w=2$) or
      become \emph{irreducible} under the joint algebra generated by relaxation and torsion,
      which would qualitatively change the accessible decompositions within
      $H^{(w)}_{\mathrm{adm}}$. :contentReference[oaicite:7]{index=7}

    \paragraph{Outlook.}
      If such irreducibility sets in for $w\ge 3$, it becomes the natural and essentially unique
      locus in the framework where a large, parameter-sparing separation between composite
      and elementary mass scales \emph{could} be forced structurally, rather than inserted
      phenomenologically.
