\subsection{Spectral Stability and the Lamb Shift}
\label{subsec:lamb-shift}

The stability of charged particle-like excitations is governed not only
by topological admissibility but also by the fine spectral structure
induced by the coupling between localized modes and the global relaxation
flow.
While the linear effective description predicts degenerate energy levels
for certain atomic configurations, this degeneracy is lifted once
non-linear saturation effects and projectability constraints are taken
into account.

In standard QED, the Lamb shift arises from radiative corrections
requiring renormalization.
In Cosmochrony, no vacuum degrees of freedom are introduced.
Atomic bound states correspond to admissible localized relaxation modes.
$S$-states ($\ell=0$) probe finer scales of the projection fiber where
Born--Infeld saturation and spectral frustration are maximal, whereas
$P$-states ($\ell=1$) remain less sensitive to these inner-core
constraints.

The effective Born--Infeld dynamics induces a finite upward shift of the
$S$-state energy:
\begin{equation}
  \Delta E_{\mathrm{Lamb}}
    \sim \kappa\, \alpha^5 m_e c^2 ,
\end{equation}
where $\alpha$ is the dimensionless ratio between the local relaxation
flux and the maximal saturation flux~$c_\chi$, and $\kappa$ is a
numerical factor of order unity.
Numerically, $\alpha^5 m_e c^2 \simeq 10^{-5}\,\mathrm{eV}$,
corresponding to a frequency shift of order $1\,\mathrm{GHz}$, in
agreement with the observed $2S_{1/2}$--$2P_{1/2}$ splitting.
This correction is intrinsically finite thanks to the maximal relaxation
speed~$c_\chi$.

\paragraph{Hyperfine Structure.}
The same spectral-probe logic extends to hyperfine structure.
When the electron occupies an $S$-state, its non-vanishing density at the
nucleus brings electronic and nuclear torsional cores into spectral
proximity.
Hyperfine splitting reflects the relative alignment of torsional fluxes:
parallel alignments accumulate frustration and oppose relaxation, while
antiparallel alignments partially compensate torsion.
Together with the Lamb shift, hyperfine structure emerges as a finite
spectral correction governed by Born--Infeld saturation at different
relational scales.
