\subsection{Relation to Quantum Formalism}
  \label{sec:relation-to-quantum-formalism}

  The formal apparatus of quantum mechanics arises as an effective
  framework organizing admissible projected descriptions in regimes where
  localization, linearity, and approximate factorization hold.
  Quantum mechanics is not replaced but reinterpreted as an effective
  theory whose validity is restricted to regimes admitting a stable,
  approximately linear, spacetime-based description.

  \paragraph{Status of the Wavefunction.}
    \label{subsec:status-of-the-wavefunction}
    The wavefunction~$\psi$ is not a fundamental physical entity but a
    statistical encoding of admissible local reprojections compatible with a
    given non-factorizable projected configuration.
    Its complex phase encodes relational constraints between alternative
    local descriptions; its modulus determines the statistical weight of
    accessible reprojections.

  \paragraph{Why a Hilbert Space.}
    \label{subsec:why-hilbert-space}
    Because the effective description has no access to~$\chi$ itself, it
    must represent, at the level of observables alone, the coexistence of
    distinct configurations belonging to the same fiber $\Pi^{-1}(o)$.
    This representational requirement imposes three successive constraints.
    First, contributions from different pre-images must be combinable:
    the space of effective descriptions must be additive.
    Second, observed interference phenomena---constructive or destructive
    superposition of projectively admissible contributions---require a phase
    structure that a purely real or purely positive representation cannot
    support: complex scalars are forced.
    Third, the effective dynamics must preserve the total statistical weight
    of admissible reprojections, imposing a conserved norm.
    Together, these three constraints single out a complex Hilbert space as
    the minimal representational closure compatible with non-injective
    projection.
    This is not a postulate but a structurally motivated consequence of
    the non-injectivity of~$\Pi$; it does not yet constitute a fully
    derived theorem, and the precise conditions under which the
    representation is unique are left for future work.

  \paragraph{Emergence of Hilbert Space Structure.}
    \label{subsec:emergence-of-hilbert-space-structure}
    Quantum superposition is the observable trace of fiber multiplicity:
    a state of the form $\alpha\,|{o_1}\rangle + \beta\,|{o_2}\rangle$
    encodes, within the effective description, the coexistence of
    projectively admissible contributions from $\Pi^{-1}(o_1)$
    and~$\Pi^{-1}(o_2)$, not a literal simultaneous occupation of two
    distinct ontic states.
    In regimes where relational constraints are weak and approximately
    factorizable, distinct admissible descriptions can be combined, giving
    rise to approximate linear structure.
    The inner product encodes mutual compatibility; orthogonality corresponds
    to mutually exclusive descriptive regimes.
    Unitary evolution is a consistency-preserving transformation valid as long
    as projectability and factorizability remain intact.
    The structural derivation of the effective Schr\"odinger dynamics from
    these constraints is given in Appendix~\ref{app:schrodinger-derivation}.

  \paragraph{Emergence of the Schr\"odinger Equation.}
    \label{subsec:schrodinger-emergence}

    In the non-relativistic regime, the effective dynamics of projected modal amplitudes
    reduces to a Schr\"odinger-type evolution equation.
    Within Cosmochrony, this equation is not taken as fundamental, but arises as the
    weak-amplitude, approximately injective limit of the projected dynamics of admissible
    $\chi$-configurations.
    Separating a rapidly varying phase from a slowly varying envelope yields
    \begin{equation}
      i\hbar_{\mathrm{eff}}\,\partial_t \psi
      = -\frac{\hbar_{\mathrm{eff}}^2}{2m_{\mathrm{eff}}}\nabla^2 \psi
      + V_{\mathrm{eff}}(x)\psi.
      \label{eq:schrodinger-effective}
    \end{equation}
    A structural derivation of this effective equation is given in
    Appendix~\ref{app:schrodinger-derivation}.

  \paragraph{Operators and Non-Commutativity.}
    \label{subsec:operators-from-chi}
    The canonical commutation relation $[\hat{x}, \hat{p}] = i \hbar_{\mathrm{eff}}$
    expresses the non-commutativity of geometric measurements induced by projection from the pre-geometric substrate.
    The uncertainty principle emerges as a geometric consistency condition on admissible effective descriptions.

    This canonical non-commutativity operates at the level of effective spacetime observables.
    It should be distinguished from the spectral and sectorial non-commutativities discussed in
    Sections~\ref{sec:noncommutativity-mass} and~\ref{subsec:flavor-structure-cp}, which arise within the projection
    fiber itself and manifest respectively as inertial mass and CP-violating geometric phases.

  \paragraph{Origin of Quantization.}
    \label{subsec:origin-of-quantization}
    Only a restricted class of localized projected configurations admits
    long-lived, internally consistent descriptions.
    Admissible configurations form discrete equivalence classes, yielding
    effective energy quantization.
    Planck's constant emerges as a universal conversion factor characterizing
    the minimal structural scale at which projected descriptions remain
    stable.

  \paragraph{Measurement and the Born Rule.}
    \label{subsec:measurement-and-the-born-rule}
    Measurement outcomes correspond to effective reprojections onto mutually
    exclusive descriptive regimes.
    The Born rule emerges as the unique stable measure on the space of
    admissible projected descriptions that remains invariant under loss of
    phase coherence, coarse-graining, and environmental coupling.

  \paragraph{Entanglement and Nonlocal Correlations.}
    \label{subsec:entanglement-nonlocality}
    Entangled systems arise when a unified relational configuration admits an
    effective projection onto spatially separated degrees of freedom.
    Correlations are fully compatible with relativistic causality
    (Appendix~\ref{subsec:non-factorization-entanglement}).

  \paragraph{Spin and Statistics.}
    \label{subsec:spin-statistics}
    Spin emerges as a topological property of admissible projected
    configurations.
    $4\pi$-periodic configurations correspond to fermionic behavior;
    $2\pi$-periodic ones to bosonic behavior.
    The spin--statistics connection follows from the same topological
    constraints
    (Appendix~\ref{subsec:4pi_soliton}).

  \paragraph{Orbital Geometry as Probabilistic Visibility.}
    \label{sec:orbital_geometry_visibility}
    Atomic orbitals correspond to effective probabilistic visibility patterns
    of admissible projected descriptions.
    Regions of high probability correspond to domains where consistent
    reprojection is most robust; nodal regions correspond to incompatible
    descriptive regimes.

  \paragraph{Scope and Limitations.}
    \label{subsec:scope-and-limitations}
    All standard quantum-mechanical formalisms remain valid and unchanged
    within their domains.
    Cosmochrony provides an interpretative and unificatory account of the
    structural origin of quantum phenomena without introducing new degrees of
    freedom or hidden variables.
    A complete formal correspondence between the relational substrate and the
    operator-based structures of quantum theory is left for future work.
