\subsection{Structural Origin of Quantum Correlations and Non-Locality}
  \label{app:structural-origin-quantum-correlations}

  The non-injective projection $\Pi$ provides a structural
  reinterpretation: what appear as two spatially separated particles may
  correspond to a single relational configuration in $\chi$.
  Entanglement is the manifestation of a single $\chi$-source through
  multiple non-injective projective images.
  Spin correlations arise from torsional conservation: a local
  stabilization at~$A$ structurally constrains admissible projective
  states at~$B$ originating from the same source.

  The mechanism through which a single $\chi$ source constrains two spatially separated
  projective images can be made more precise.
  Let $I(\omega)$ denote a global invariant of the configuration $\omega \in \Omega_\chi$---a
  conserved topological quantity such as total winding number, total charge, or total angular
  momentum.
  Such an invariant belongs to a sector of the substrate that is not individually projectable:
  it cannot be decomposed as $I_A \oplus I_B$ with $I_A$ and $I_B$ each independently
  representable as local effective observables in $O_A$ and $O_B$.
  It follows that the fiber $\Pi^{-1}(o_{AB})$ cannot be factored as
  $\Pi^{-1}(o_A) \times \Pi^{-1}(o_B)$: any admissible description must treat
  $I(\omega)$ as a single global property of the joint projected class.
  The two observers at $A$ and $B$ do not measure two independent properties of the substrate.
  They read the same underlying invariant from two distinct local relational frames
  $\Pi_A$ and $\Pi_B$.
  The perfect correlations that characterize entangled states reflect the fact that there is only
  one invariant to be distributed across the two frames, not two independent quantities
  that happen to agree.
  Bell-type factorization fails because it assumes that the effective variable
  $\lambda = \Pi(\omega)$ admits a product representation
  $\Pi^{-1}(o_A) \times \Pi^{-1}(o_B)$; the infra-projectable
  character of $I(\omega)$ makes this assumption structurally inadmissible.

  The geometric origin of the $\cos\theta$ correlation law and the Tsirelson bound
  can be stated compactly in terms of the substrate invariant structure.
  Let the substrate invariant associated with a singlet configuration be represented by
  a unit element of $S^3$, reflecting the $\mathrm{SU}(2)$ structure of the spin sector
  of the projection fiber.
  Directional measurements correspond to projections of this $S^3$ invariant onto axes
  in the effective space $S^2$, with spin observables satisfying
  $(\boldsymbol{\sigma}\cdot\mathbf{a})^2 = 1$.
  The resulting mean correlation is
  $E(\mathbf{a}, \mathbf{b}) = \langle(\boldsymbol{\sigma}\cdot\mathbf{a})
  \otimes (\boldsymbol{\sigma}\cdot\mathbf{b})\rangle = -\mathbf{a}\cdot\mathbf{b}$,
  reproducing the quantum mechanical prediction.
  Note that a classical model with binary outcomes
  $A = \mathrm{sign}(I\cdot\mathbf{a})$, $B = -\mathrm{sign}(I\cdot\mathbf{b})$
  and $I$ uniformly distributed on $S^2$ yields $E = 1 - 2\theta/\pi$, not
  $-\cos\theta$: the correct $\cos\theta$ law requires the $\mathrm{SU}(2)$ operator
  structure, not a classical sign projection.
  The CHSH combination satisfies $|S| \leq 2\sqrt{2}$ as a consequence of the
  Cauchy--Schwarz inequality applied to unit vectors in $\mathbb{R}^2$; this bound
  is saturated at the optimal orientation angles and coincides with the Tsirelson
  bound.
  Both results---the $\cos\theta$ law and the bound $2\sqrt{2}$---are therefore
  structural consequences of the spherical geometry of the projection space induced
  by $\Pi$, not independent quantum postulates.
  A factorizable hidden-variable model would require $\Pi^{-1}(o_{AB}) =
\Pi^{-1}(o_A) \times \Pi^{-1}(o_B)$, which is incompatible with the global
  uniqueness of $I(\omega)$; Bell factorization fails for the same reason that the
  invariant cannot be split into two independently projectable local components.

  The framework remains ontologically realist at the substrate level but
  structurally non-local with respect to the emergent spacetime.
