\subsection{Non-Injectivity as a Structural Necessity of Genuine Emergence}
  \label{subsec:non-injectivity-not-ad-hoc}

  Non-injectivity of the projection $\Pi : \Omega \to \mathcal{O}$ is not a modelling
  choice introduced to generate desired phenomenology.
  It is a structural necessity of any genuinely emergent physical description, provably
  so under minimal hypotheses~\cite{Beau2026l}.

  The core result is the following.
  Let $\Pi : \Omega \to \mathcal{O}$ be surjective, with observables entirely defined
  by $\Pi$ --- meaning that $\mathcal{O}$ carries no independent microstate-resolving
  structure beyond what $\Pi$ provides --- and suppose that $\mathcal{O}$ is not
  structurally isomorphic to $\Omega$.
  Then $\Pi$ is necessarily non-injective and the associated projection entropy
  satisfies $S_\Pi > 0$.
  The proof is elementary: if $\Pi$ were injective, it would be a structural isomorphism
  $\mathcal{O} \cong \Omega$, contradicting the assumption of genuine emergence.
  A fully injective emergent description is therefore a contradiction in terms ---
  it reduces emergence to a mere relabelling of the fundamental level.

  Three structural consequences are co-extensive with this necessity.

  \begin{enumerate}
    \item \textbf{Descriptive redundancy.}
    The multiplicity of pre-images $\Pi^{-1}(o)$ for a given $o \in \mathcal{O}$
    defines a structural projection entropy
    \begin{equation}
      S_\Pi \;\equiv\; -\sum_{o \in \mathcal{O}} \nu(o)\,\log\nu(o),
      \qquad \nu(o) = \mu\!\left(\Pi^{-1}(o)\right),
    \end{equation}
    which is strictly positive whenever $\Pi$ is non-injective and measures an
    objective structural property of $\Pi$, not epistemic ignorance about the
    underlying configuration.
    In any genuinely emergent framework, descriptive redundancy, non-injectivity
    of $\Pi$, and $S_\Pi > 0$ are co-extensive~\cite{Beau2026l}.

    \item \textbf{Emergent gauge structure.}
    If $\Pi$ were injective and compatible with relational locality, parallel
    transport of effective states along any closed loop in $\mathcal{O}$ would lift
    uniquely to a closed loop in $\Omega$, and the holonomy group of the induced
    fiber bundle would reduce to the identity.
    Trivial holonomy precludes any non-trivial fiber re-identification symmetry and
    therefore any emergent gauge degree of freedom (Appendix~A.9).
    Any non-trivial gauge structure therefore requires $\Pi$ to be non-injective,
    and hence $S_\Pi > 0$.

    \item \textbf{Projective Bell-violation.}
    Non-injectivity implies that the joint outcome distribution of spatially separated
    measurements cannot be factorized into independent local contributions, because
    the underlying configurations are not individually assigned to localized observable
    outcomes.
    Bell-type correlations arise as the minimal observable signature of $S_\Pi >
    0$~\cite{Beau2026b}.
  \end{enumerate}

  In the Cosmochrony framework, the intermediate geometric hypothesis required to
  pass from the universal necessity of non-injectivity to the specific gauge group
  $\mathrm{SU}(2) \times U(1)$ is explicitly realised: the projection
  $\Pi : \mathcal{C}_\chi \to M$ defines a principal bundle with fibre $S^3$, and
  the Born--Infeld saturation bound selects the unique admissible infrared dynamics.
  The route from non-injectivity to $\mathrm{SU}(2)$ via Born--Infeld parity, Weil
  representations, and the Hurwitz classification is detailed in Appendix~\cite{sec:appendix-math} and the
  O-series companion papers~\cite{Beau2026a18,Beau2026a19}.

  In summary, non-injectivity is the unique structural condition that simultaneously
  accounts for descriptive redundancy ($S_\Pi > 0$), emergent gauge structure
  (non-trivial holonomy), finite transport capacity (bounded effective flux), and
  projective Bell-violation (loss of probabilistic factorization).
  It is a theorem, not an assumption.
