\subsection{Cosmic Microwave Background}
  \label{subsec:cmb}

  Within the Cosmochrony framework, the cosmic microwave background (CMB) does not encode primordial fluctuations
  generated during a distinct inflationary phase.
  It reflects instead the imprint of early relaxation and reprojection processes as the relational $\chi$ substrate
  transitioned toward a regime admitting stable geometric descriptions.

  At this stage, the universe was not yet structured by well-defined spatial separation or causal horizons.
  Large-scale correlations observed in the CMB therefore arise naturally from the global relational coherence of $\chi$
  prior to geometric differentiation, rather than from superluminal expansion within spacetime.

  The acoustic features observed in the temperature power spectrum admit an effective interpretation
  as resonance patterns arising once a stable geometric regime is reached, where the standard photon--baryon
  plasma description becomes operationally meaningful.
  In Cosmochrony, the distinctive imprint of $\chi$ enters primarily through the primordial conditions and
  large-scale admissibility constraints, rather than by replacing the established microphysics of the plasma
  itself~\cite{SachsWolfe1967,HuWhite1997}.

  In this perspective, the CMB encodes a ``fossil record'' of the emergence of spacetime itself.
  Its large-angle correlations and statistical properties are consequences of the pre-geometric relational structure of
  $\chi$, inherited by the emergent spacetime description without requiring an inflationary epoch or superluminal causal
  processes~\cite{Planck2020}.

  \paragraph{Effective closure: primordial spectrum as the interface to CMB pipelines.}
    At the level of quantitative prediction, we adopt an effective closure in which the standard
    Boltzmann transfer functions remain unchanged once a stable geometric description has emerged.
    Within this approximation, Cosmochrony modifies the CMB angular spectra only through the primordial
    curvature spectrum $P_\zeta(k)$:
    \begin{equation}
      C_\ell^{XY}
      =
      4\pi \int_0^\infty \frac{dk}{k}\, P_\zeta(k)\,
      \Delta_\ell^{X}(k)\,\Delta_\ell^{Y}(k),
    \end{equation}
    where $\Delta_\ell^{X}(k)$ are the usual transfer functions for temperature and polarization.
    The Cosmochrony imprint is encoded in a projectability-induced infrared filter and an emergent tilt,
    \begin{equation}
      P_\zeta(k)
      =
      A_s\left(\frac{k}{k_*}\right)^{n_\chi-1}
      C^2\!\left(\frac{k}{k_p}\right),
      \qquad
      C(x)\xrightarrow{x\ll 1}0,\quad C(x)\xrightarrow{x\gg 1}1,
    \end{equation}
    where $k_p$ is the effective cutoff scale associated with the freeze-out of admissible projected
    descriptions.
    This construction preserves the successful peak structure of the standard plasma transfer physics,
    while allowing Cosmochrony to address large-angle anomalies through a controlled modification of the
    primordial baseline.

  This suppression may explain the observed $\ell=2$ anomaly in Planck data, without invoking inflationary mechanisms.

  Because the emergence of geometric descriptions and the stabilization of projection
  modes are governed by the same underlying spectral structure of the $\chi$ substrate,
  fundamental ratios characterizing stable configurations at the micro-scale may leave
  scale-independent imprints on cosmological observables.

  \paragraph{Relational spectral scales and angular multipoles.}
    The admissibility cutoff and large-angle modulations can be expressed directly in terms of the
    low-lying spectrum of the relational operator $L_\chi$ introduced in
    Section~\ref{sec:relational_projection}.
    Once a stable geometric regime is reached, the long-wavelength sector admits an effective Laplacian
    interpretation, yielding the canonical identification
    \begin{equation}
      k_n \;\propto\; \sqrt{\lambda_n(\lambda_*)},
      \qquad
      \ell_n \;\approx\; k_n\,D_A(z_*),
    \end{equation}
    and therefore the robust ratio constraint
    \begin{equation}
      \frac{\ell_n}{\ell_1}
      \;\approx\;
      \sqrt{\frac{\lambda_n}{\lambda_1}}.
      \label{eq:ell_ratio_spectral}
    \end{equation}
    Importantly, these $\ell_n$ characterize \emph{relational angular modulations} associated with
    admissibility and freeze-out, and are not assumed to coincide with the acoustic peak indices of the
    photon--baryon plasma.

    \begin{remark}[Spectral ratio $\lambda_2/\lambda_1=8/3$ and angular-scale constraints]
      Numerical relaxation experiments reported in Appendix~D.7 indicate that, across a broad class of
      admissible configurations, the low-lying spectrum of $L_\chi$ exhibits the stable ratio
      \begin{equation}
        \frac{\lambda_2}{\lambda_1} \;\simeq\; \frac{8}{3}.
      \end{equation}
      If this ratio persists in the continuum limit, Equation~\eqref{eq:ell_ratio_spectral} implies the
      characteristic constraint
      \begin{equation}
        \frac{\ell_2}{\ell_1} \;\approx\; \sqrt{\frac{8}{3}} \;\simeq\; 1.63,
      \end{equation}
      which can be searched for as a scale-independent modulation signature in the low-$\ell$ sector.
    \end{remark}

  \paragraph{Polarization and tensor modes (outlook).}
    Possible signatures of the fundamental spectral ratio $\lambda_2/\lambda_1 = 8/3$
    in CMB polarization observables, including constraints on tensor modes, are discussed separately in
    Section~\ref{subsec:cmb-polarization-signatures} as testable predictions of the framework.

  \paragraph{Clarification: relation to acoustic peaks.}
    In standard cosmology, the acoustic peaks arise from oscillations of the photon--baryon fluid
    prior to recombination, and their detailed positions and heights are controlled by well-tested
    plasma microphysics.
    The relational eigenvalues $\{\lambda_n\}$ of $L_\chi$ are \emph{not} fluid oscillation modes.
    They characterize the spectral geometry of admissible $\chi$ configurations at freeze-out and thus
    constrain the large-angle primordial baseline and admissibility-induced modulations.
    Accordingly, the present work does not identify $\ell_n \propto \sqrt{\lambda_n}$ with the acoustic
    peak indices; instead, the acoustic peak structure is retained through standard transfer functions,
    while Cosmochrony enters through $P_\zeta(k)$ and low-$\ell$ effects.
