\subsection{Relation to Bell Inequalities}
  \label{subsec:relation-to-bell-inequalities}

  Bell's theorem~\cite{Bell1964} rules out any ontological completion of
  quantum mechanics based on factorizable hidden-variable models.
  Cosmochrony fully accepts Bell's theorem and identifies the ontological
  assumption that fails: the factorization hypothesis $P(a,b|x,y,\lambda) = P(a|x,\lambda)\,P(b|y,\lambda)$.

  Because the projection~$\Pi$ is generically non-injective, admissible
  projected states are not associated with independent ontic pre-images for their subsystems.
  The factorization hypothesis is not merely violated but \emph{ill-defined}
  within the space of admissible projected descriptions.

  Bell-type violations arise in an intermediate regime of projective
  compression: the effective description is sufficiently coarse-grained to
  permit subsystem separation, yet retains enough global relational structure to prevent factorization.
  In the limit of extreme coarse-graining, projected descriptions become effectively classical.

  \begin{figure}[t]
    \centering
    \begin{tikzpicture}[
      scale=1.05,
      axis/.style={->, thick},
      curve/.style={thick, smooth, blue},
      dashedline/.style={dashed, gray},
      label/.style={font=\small}
    ]
    \draw[axis] (0,0) -- (10,0)
    node[right,label] {Projection compression};
    \draw[axis] (0,0) -- (0,5)
    node[above,label] {Non-factorizable corr.};
    \draw[dashedline] (0,1.2) -- (10,1.2)
    node[right,label] {$S=2$};
    \draw[dashedline] (0,4) -- (10,4)
    node[right,label] {$S=2\sqrt{2}$};
    \draw[curve]
    plot[domain=0.5:9.5,samples=200]
    (\x,{4*exp(-0.18*(\x-5)^2)});
    \node[label, align=center] at (2.2,4.4)
      {Under-\\compression};
    \node[label, align=center] at (5,4.7)
      {Critical\\(quantum)};
    \node[label, align=center] at (8,4.4)
      {Over-\\compression};
    \draw[dashedline] (2.2,0) -- (2.2,4.2);
    \draw[dashedline] (5,0) -- (5,4.6);
    \draw[dashedline] (8,0) -- (8,4.2);
    \end{tikzpicture}
    \caption{Bell inequality violations as a function of the
    compression induced by~$\Pi$.
    Non-factorizable correlations emerge in an intermediate regime.}
    \label{fig:bell-compression-regime}
  \end{figure}

  The dependence of entanglement correlations on measurement orientation acquires a
  natural reading within this framework.
  Each measurement apparatus defines a local relational frame, corresponding to a
  directional projection operator $\Pi_{\mathbf{a}}$ or $\Pi_{\mathbf{b}}$ parametrized
  by the orientation vectors $\mathbf{a}$ and $\mathbf{b}$.
  The local outcomes $o_A = \Pi_{\mathbf{a}}(\omega)$ and $o_B = \Pi_{\mathbf{b}}(\omega)$
  are therefore two readings of the same infra-projectable invariant $I(\omega)$ from
  two distinct relational frames.
  For a spin-singlet configuration ($S_\mathrm{tot} = 0$), the infra-projectable invariant
  is represented by a unit element of $S^3$, reflecting the $\mathrm{SU}(2)$ structure of
  spin degrees of freedom in the projection fiber (Section~5.2).
  Directional measurements correspond to projections of this $S^3$ invariant onto axes
  in the effective space $S^2$; the spin observables satisfy $(\boldsymbol{\sigma}\cdot\mathbf{a})^2 = 1$
  and the mean correlation between two measurement directions is
  \begin{equation}
    E(\mathbf{a}, \mathbf{b})
    = \langle (\boldsymbol{\sigma}\cdot\mathbf{a}) \otimes (\boldsymbol{\sigma}\cdot\mathbf{b}) \rangle
    = -\mathbf{a} \cdot \mathbf{b} = -\cos\theta,
    \label{eq:Bell-cosine}
  \end{equation}
  where $\theta = \angle(\mathbf{a},\mathbf{b})$.
  This result follows from the $\mathrm{SU}(2)$ geometry of the projection space: the
  $\cos\theta$ law is not derived from a classical sign model---which would yield
  $1 - 2\theta/\pi$ and not saturate the Tsirelson bound---but reflects the
  non-commutative structure of directional projections of an $S^3$ invariant.
  No dynamical influence between $A$ and $B$ is required; the $\cos\theta$ law
  is a structural consequence of the $\mathrm{SU}(2)$ geometry of $\Pi$.

  The maximal violation of the CHSH inequality follows from the same geometric
  structure.
  The CHSH combination
  $S = E(\mathbf{a},\mathbf{b}) + E(\mathbf{a},\mathbf{b}') +
  E(\mathbf{a}',\mathbf{b}) - E(\mathbf{a}',\mathbf{b}')$
  is a linear combination of scalar products of unit vectors in $\mathbb{R}^2$.
  The identity
  \begin{equation}
    \bigl|\mathbf{a}\cdot\mathbf{b} + \mathbf{a}\cdot\mathbf{b}'
    + \mathbf{a}'\cdot\mathbf{b} - \mathbf{a}'\cdot\mathbf{b}'\bigr|
    \;\leq\; 2\sqrt{2}
    \label{eq:Tsirelson-geom}
  \end{equation}
  holds for all unit vectors and is saturated by the optimal configuration
  $\mathbf{a} = (1,0)$, $\mathbf{a}' = (0,1)$,
  $\mathbf{b} = \tfrac{1}{\sqrt{2}}(1,1)$, $\mathbf{b}' = \tfrac{1}{\sqrt{2}}(1,-1)$,
  which yields $|S| = 2\sqrt{2}$.
  The Tsirelson bound is therefore not a postulate of quantum theory imported into
  the framework: it is a geometric constraint on the space of directional projections
  of a single substrate invariant.
  Its value $2\sqrt{2}$ reflects the Hilbert-space geometry associated with the
  $\mathrm{SU}(2)$ structure of the projection fiber (Section~5.2).
  The classical bound $|S| \leq 2$ corresponds to the regime in which the
  projection becomes effectively injective and the invariant can be assigned
  independent local pre-images, restoring factorizability.

  No hidden variables and no superluminal influence are introduced.
  Quantum nonlocality is ontological rather than dynamical: the failure of
  Bell-type factorizability arises from the structure of admissible projected descriptions, not from nonlocal
  interaction.
