% ----------------------------------------------------------------------------
% Section 6.3 --- Emergent Curvature
% From former ยง7.3, condensed
% ----------------------------------------------------------------------------
\subsection{Emergent Curvature}
\label{subsec:emergent-curvature}

Spatial variations in admissible relaxation ordering lead to non-uniform
correlation patterns within projected descriptions.
When a smooth geometric parametrization is applicable, these
non-uniformities are summarized by gradients of an emergent metric
structure, reproducing the phenomenology traditionally attributed to
curved spacetime: gravitational time dilation, geodesic deviation, and
lensing effects.

\paragraph{Einstein's Equations as a Structural Equilibrium Principle.}
\label{subsec:einstein-equilibrium-principle}
When projected configurations admit a smooth, locally injective geometric
description, their collective structural constraints can be summarized by
an effective metric~$g_{\mu\nu}$.
Einstein's equations emerge \emph{necessarily} as the unique consistency
condition relating curvature to the effective distribution of
relaxation-resistant configurations.
The Einstein tensor encodes a geometric identity constraining admissible
macroscopic descriptions; the stress--energy tensor summarizes how
localized configurations resist relaxation ordering.
Einstein's equations therefore express a balance condition between
geometry and physical structure.

This interpretation explains the extraordinary universality of general
relativity: as long as a smooth spacetime description exists and
relaxation ordering is monotonic, bounded, and weakly inhomogeneous, the
same geometric relations must hold, independently of the microscopic
nature of the underlying substrate.
General relativity thus plays a role analogous to thermodynamics---exact
within its domain, silent outside it.

When projection ceases to be locally injective, Einstein's equations do
not break down; they simply no longer apply, because the concept of
spacetime has not yet emerged.
