\subsection{Relation to Bell Inequalities}
  \label{subsec:relation-to-bell-inequalities}

  Bell's theorem~\cite{Bell1964} establishes that no physical theory reproducing the full
  set of quantum mechanical predictions can simultaneously satisfy locality,
  realism understood as outcome determinism conditioned on hidden variables, and statistical independence.
  The empirical violations of Bell inequalities therefore rule out any ontological
  completion of quantum mechanics based on factorizable hidden-variable models.

  The Cosmochrony framework fully accepts Bell's theorem as a fundamental constraint.
  It does not seek to evade or weaken Bell inequalities, nor to restore locality or realism in their classical sense.
  Instead, it identifies the precise ontological assumption that fails within
  Bell-type derivations when applied to admissible projected descriptions.

  \paragraph{Failure of ontological factorisability.}
    Standard Bell-type arguments assume that joint outcome probabilities admit a factorization of the form
    \begin{equation}
      P(a,b|x,y,\lambda) = P(a|x,\lambda)\,P(b|y,\lambda),
    \end{equation}
    where $\lambda$ denotes a complete specification of the underlying ontic state.
    In Cosmochrony, such a decomposition is not available, not because of nonlocal
    dynamical influences, but because admissible effective descriptions arise from
    \emph{non-injective projections} of the underlying $\chi$ substrate.

    Distinct fundamental configurations of $\chi$ may correspond to the same effective
    projected description, while remaining globally constrained.
    As a consequence, entangled systems correspond to equivalence classes of
    non-separable configurations that do not admit independent ontic pre-images for their subsystems.
    There exists no hidden variable $\lambda$ with respect to which conditional
    factorization could be restored, even in principle.

  \paragraph{No hidden variables and no superluminal influence.}
    The non-injectivity of the projection does not introduce hidden variables, local or nonlocal.
    The underlying degrees of freedom are neither accessible nor conditionable,
    and they cannot be used to screen correlations or define outcome-determining parameters.
    Accordingly, the violation of Bell inequalities in Cosmochrony does not rely on
    superluminal signalling, retrocausality, or a breakdown of relativistic causal structure.

    Correlations are not mediated, transmitted, or dynamically generated between spatially separated subsystems.
    They reflect global admissibility constraints on projected descriptions, inherited
    from the holistic structure of the underlying configuration space.

  \paragraph{Ontological rather than dynamical nonlocality.}
    In this sense, quantum nonlocality in Cosmochrony is ontological rather than dynamical.
    The failure of factorization arises from the structure of the projection map itself,
    not from any nonlocal interaction or exchange of information.
    Bell-type violations are therefore understood as a manifestation of the
    non-separable nature of admissible descriptions, rather than as evidence for superluminal influences.

    This interpretation preserves the full empirical content of quantum mechanics,
    respects relativistic causality at the operational level, and provides a structural
    explanation for the inevitability of entanglement correlations within a non-injective relational framework.
