\subsection{The Geometry of the
\texorpdfstring{$\Pi$}{Π} Subspace}
  \label{sec:geometry-pi}

  The relational substrate~$\chi$ is not accessed in its full structural
  complexity but through admissible local projections onto a reduced
  projection fiber $\Pi \cong S^3$.
  Here $\Pi$ denotes the projection \emph{fiber}, i.e.\ the space of
  admissible internal representatives associated with a projected
  configuration, not the projection map itself.

  The identification $\Pi \cong S^3$ is not imposed \emph{a priori} but
  follows from the minimal compact geometry (in dimension and connectivity)
  required to support non-injective yet stable projections.
  Its associated Hopf fibration naturally induces an effective
  $SU(2)\times U(1)$ symmetry structure, which emerges as an invariance of
  the projection process.

  The metric on~$\Pi$ is dynamically induced by the local density of
  relational connections encoded in the underlying relational graph~$G$,
  used as a numerical and representational support, not as a physical
  discretization.
  The mapping from the global graph Laplacian~$\Delta_G$ to the effective
  projected Laplacian~$\Delta_\Pi$ is
  \begin{equation}
    \Delta_\Pi = P^\dagger \Delta_G P ,
  \end{equation}
  where $P$ denotes the projection operator onto the admissible spectral
  subspace.
  The smooth three-sphere geometry corresponds to a large-$N$ spectral
  coarse-graining limit of the discrete relational structure.

  \paragraph{Throughput constraints and channel counting.}
    The projection $\Pi$ does not merely discard information.
    It enforces an admissibility constraint: local updates must remain operationally compatible with the surrounding
    global projected state~$U$.
    This constrains the total admissible throughput by the number of independent relational channels (fibers) that can
    be synchronized across the update interface within one refresh step.
    Accordingly, capacity scalings commonly expressed as ``area-law'' behavior should be understood as consequences of
    projective throughput constraints rather than geometric postulates.
    This point becomes central when discussing resolvability bounds across scales in
    Section~\ref{subsec:projective-resolvability-bound}.
