\subsection{Topological Invariants of the Projection Fiber}
  \label{sec:topological-constraints}

  The stability of localized physical descriptions within~$\Pi$ is governed
  by the conservation of topological invariants.
  When an excitation of~$\chi$ admits a closed, non-contractible
  configuration within~$\Pi$, it forms a persistent topological obstruction
  to complete relaxation.
  Such configurations cannot be eliminated by smooth deformation without
  violating admissibility constraints.
  These obstructions are described in terms of knot-like (non-contractible loop and
  self-linking) topological structures within the projection fiber.
  Once formed, they impose global constraints on the relaxation process and
  give rise to long-lived, localized projected configurations.
  Particles correspond to stable topological defects of the projection.

  The homotopy groups of the admissible fiber~$\Pi$ and of its admissible
  base determine which topological invariants are structurally available,
  and hence which physical phenomena can arise.
  Three consequences are established on the canonical admissible configuration model;
  their extension to the full effective configuration space~$\mathcal{C}_{\mathrm{eff}}$
  remains open (see~\cite{Beau2026k} for the complete treatment).

  \subsubsection*{Charge quantisation as a $\pi_1$ invariant}
    \label{sec:charge-quantisation-pi1}

    The Hopf fibration of the projection fiber $\Pi \cong S^3$,
    \begin{equation}
      S^1 \hookrightarrow S^3 \xrightarrow{\pi_H} S^2,
    \end{equation}
    decomposes~$\Pi$ into a circular fiber~$S^1$ and a two-dimensional base~$S^2$.
    The fundamental group of the $U(1)$ fiber is
    \begin{equation}
      \pi_1(S^1) \cong \mathbb{Z}.
    \end{equation}
    A closed loop in the fiber defines a winding number $w \in \mathbb{Z}$,
    which is a topological invariant: it cannot change under continuous deformation
    of an admissible configuration.

    \begin{proposition}[Charge quantisation]
      \label{prop:charge-quantisation}
      In the Cosmochrony framework, electric charge is quantised as a consequence of
      $\pi_1(S^1) \cong \mathbb{Z}$.
      The admissibility constraint removes configurations exceeding the flux bound
      but does not introduce identifications that would contract the fiber,
      so the fundamental group of the fiber sector remains that of $U(1)$.
      Electric charge is therefore an integer multiple of the elementary charge~$e$,
      \begin{equation}
        q_{\mathrm{eff}} = w \cdot e, \qquad w \in \mathbb{Z}.
      \end{equation}
    \end{proposition}

    The winding number~$w$ labels distinct equivalence classes of admissible
    projected configurations and plays a central role in the organization of
    the mass spectrum (Section~\ref{sec:spectral-mass-hierarchy}).
    The energetic cost required to maintain a non-trivial winding against
    the global pressure of relaxation is perceived, at the effective level,
    as rest energy~($mc^2$).
    Mass emerges as a spectral and topological consequence of persistent
    non-contractible structures within the projection fiber.

    Unlike the standard Dirac argument~--- which derives charge quantisation from
    the existence of magnetic monopoles via single-valuedness of the wave function
    on a surrounding sphere~--- charge quantisation in Cosmochrony follows from
    the topology of the $U(1)$ fiber itself, independently of whether
    magnetic monopoles exist.
    The two mechanisms are logically independent.

  \subsubsection*{Absence of magnetic monopoles}
    \label{sec:no-monopoles}

    In the unconstrained Hopf fibration, the base manifold is the full two-sphere~$S^2$,
    with $\pi_2(S^2) \cong \mathbb{Z}$.
    A non-trivial element of $\pi_2(S^2)$ corresponds, in gauge-theoretic language,
    to a Dirac magnetic monopole: a point-like source of magnetic flux classified by
    an integer topological charge.
    If the full~$S^2$ were the admissible base, magnetic monopoles would be
    topologically admissible.

    The bounded-flux constraint $|\partial_t \chi_v| \le c_\chi$ restricts the
    set of admissible fiber configurations.
    On the canonical admissible configuration model, the constraint removes a
    neighbourhood of each pole of~$S^2$ (the points where the fiber degenerates,
    corresponding to maximal flux concentration).
    The admissible base is then
    \begin{equation}
      S^2_{\mathrm{adm}} = S^2 \setminus (D^2_N \cup D^2_S) \simeq S^1 \times (0,1),
    \end{equation}
    where $D^2_N$ and $D^2_S$ are open discs around the north and south poles.

    \begin{proposition}[Homotopy of the canonical admissible base]
      \label{prop:pi2-trivial}
      The canonical admissible base $S^2_{\mathrm{adm}}$ is homotopy equivalent
      to $S^1 \times (0,1)$.
      Its homotopy groups are
      \begin{equation}
        \pi_1(S^2_{\mathrm{adm}}) \cong \mathbb{Z}, \qquad \pi_2(S^2_{\mathrm{adm}}) = 0.
      \end{equation}
    \end{proposition}

    \begin{corollary}[No isolated magnetic charge]
      \label{cor:no-monopoles}
      On the canonical admissible model, no isolated magnetic monopole is a
      topologically admissible excitation.
    \end{corollary}

    In the standard Dirac~--~'t~Hooft~--~Polyakov framework, monopoles are
    topologically admissible and their absence must be explained dynamically.
    In Cosmochrony, the situation is reversed: the absence of monopoles is
    not a dynamical accident but a topological consequence of admissibility.
    The bounded-flux constraint removes the topological support for
    monopole-type excitations.
    Magnetic flux lines (corresponding to $\mathbf{B} = \nabla \times \mathbf{A}_T$)
    remain admissible as long as they form closed loops; only isolated magnetic
    sources are excluded.

    The extension of Corollary~\ref{cor:no-monopoles} from the canonical
    admissible base to the full effective configuration space~$\mathcal{C}_{\mathrm{eff}}$
    is well-posed but remains open.
    The expectation that the result extends is motivated by the fact that the
    bounded-flux constraint removes precisely the gradient singularities
    required to support non-trivial $\pi_2$ classes.

  \subsubsection*{Projective chirality and parity violation}
    \label{sec:projective-chirality}

    The winding number $w \in \mathbb{Z}$ carries a sign determined by the
    orientation of the fiber phase.
    In a parity-symmetric theory, configurations with $w$ and $-w$ would be
    related by a mirror transformation and have equal dynamical weight.
    In Cosmochrony, the projection $\Pi : \chi \to \chi_{\mathrm{eff}}$ is
    non-injective, and this non-injectivity can be chiral.

    \begin{definition}[Projective chirality]
      \label{def:projective-chirality}
      A projection is \emph{projectively chiral} if the effective
      stabilisation rates satisfy
      \begin{equation}
        \Gamma(Q) \neq \Gamma(-Q),
      \end{equation}
      where $\Gamma(Q)$ is the effective stabilisation rate for configurations
      of topological charge~$Q$, and $-Q$ denotes the charge-conjugate
      configuration.
    \end{definition}

    When the projection is chiral, parity is violated: mirror-related
    configurations are not equally admissible.
    Two structural mechanisms within the framework produce chirality selection.

    \paragraph{Spinorial selection.}
      The spectral admissibility programme establishes that the bounded-flux
      constraint selects spinorial ($\mathrm{spin}$-$\tfrac{1}{2}$) representations
      as the most admissible along the chain $Q_8 \subset 2I \subset SU(2)$
      (Section~\ref{sec:spectral-admissibility} and~\cite{Beau2026a1}).
      Spinorial representations carry an intrinsic handedness through the
      double-cover structure $SU(2) \to SO(3)$: the central $\mathbb{Z}_2$
      retained by binary covers is precisely the element that distinguishes
      left- and right-handed spinors.

    \paragraph{Non-injective projection.}
      The non-injectivity of the projection creates a fiber degeneracy:
      multiple $\chi$-configurations map to the same effective state.
      The orientation of the winding within this fiber is not preserved by a
      generic projection, so parity invariance is not automatic.
      A chiral projection selects one orientation over its conjugate, producing
      $\Gamma(Q) \neq \Gamma(-Q)$ without any explicit parity-breaking term at
      the substrate level.

      The maximal parity violation observed in weak interactions (polarised beta
      decay of ${}^{60}\mathrm{Co}$) can be interpreted as the manifestation of a
      maximally chiral projection in the $SU(2)$ sector of~$\Pi$.
      This correspondence is currently qualitative; a derivation of the full
      $V{-}A$ structure requires showing that the admissibility filter selects a
      single chirality sector completely, which remains open.
      The relation between projective chirality and CP asymmetry is further
      developed in Appendix~\ref{subsec:cp-asymmetry-and-chiral-selection}.

  \subsubsection*{Unified topological picture}
    \label{sec:topological-summary}

    The three results above are unified by a single structural principle: the
    effective physics is determined by the topological invariants that the
    admissibility constraint renders available, and by the orientation that the
    projection selects among their generators.

    \begin{table}[h]
      \centering
      \begin{tabular}{lll}
        \hline
        Topological invariant & Physical consequence & Mechanism \\
        \hline
        $\pi_1(S^1) \cong \mathbb{Z}$   & Charge quantisation   & Winding of $U(1)$ fiber \\
        $\pi_2(S^2_{\mathrm{adm}}) = 0$ & No magnetic monopoles & Bounded-flux reduction of base \\
        Orientation of $\pi_1$          & Parity violation      & Projective chirality \\
        \hline
      \end{tabular}
      \caption{Topological invariants of the canonical admissible configuration model
      and their physical consequences.
      The first two rows are proved on the canonical model; their extension
      to~$\mathcal{C}_{\mathrm{eff}}$ is open.
      The third row is established at the framework level as a structural mechanism.
      See~\cite{Beau2026k} for the complete treatment.}
      \label{tab:topological-invariants}
    \end{table}
