\subsection{Structural Predictions}
  \label{subsec:structural-predictions-fluctuations}

  Two structural predictions follow from the projective picture of fluctuations developed
  in Section~\ref{subsec:projective-origin-of-fluctuations}.

  First, in a regime where the effective projection becomes \emph{locally} injective,
  the local projective entropy $S_\Pi$ is suppressed and quantum fluctuations are
  correspondingly reduced.
  This suppression has a concrete realisation in the spectral admissibility programme:
  in~Q9~\cite{Beau2026q9}, the position sector of the effective Dirichlet form is shown
  to vanish in the large-$q$ limit (Q9 Lemma~4.3: $\lim_{q\to\infty}\mathcal{E}^X_q(f,f) = 0$),
  demonstrating that intra-fibre degrees of freedom are expelled from
  $\mathcal{O} = \mathrm{Im}\,\Pi$ as admissibility is imposed.
  Since the ordering differences between co-projecting configurations are precisely
  what generate distinguishable successive effective states, a locally injective regime
  implies $U_{n+1} = U_n$ in that region: the projected description becomes locally
  stationary and operational time ceases to advance there.
  This is the projective reading of decoherence and of the classical limit: it does not
  require environmental averaging as a primitive assumption, but follows from the
  local reduction of the co-projecting fiber.
  Such locally injective regimes are admissible; they correspond to what is ordinarily
  described as the quasi-classical regime.
  A locally injective regime is one in which the restriction of~$\Pi$ to the accessible
  configuration neighborhood admits a single dominant element of~$\Pi^{-1}(o)$:
  distinct substrate configurations within that neighborhood produce distinguishable
  projected states, even though the global projection remains non-injective.

  Second, a \emph{globally} injective projection---one in which $|\Pi^{-1}(o)| = 1$
  for all observables simultaneously---would eliminate all projective entropy and
  terminate temporal ordering everywhere.
  This is structurally excluded on two independent grounds.
  The configuration space $\Omega_\chi$ of the substrate is structurally richer than the
  observable space $O$: the projection $\Pi$ is an irreducible coarse-graining, so
  $|\Omega_\chi| \gg |O|$ and a globally injective $\Pi$ is mathematically incompatible
  with the compression implicit in any effective description.
  Furthermore, a globally stationary projection would imply $U_{n+1} = U_n$ for all
  effective observers simultaneously, which is the cosmological counterpart of absolute
  zero and is excluded by the same argument developed in Section~\ref{subsec:absolute-zero-necessity}.
  The persistence of a minimal global non-injectivity is therefore not merely compatible
  with temporal evolution; it is a structural requirement for it.
