% ----------------------------------------------------------------------------
% Section 3.2 --- Hamiltonian Derivation of the Evolution Equation
% From former ยง5.2
% ----------------------------------------------------------------------------
\subsection{Hamiltonian Derivation of the Evolution Equation}
\label{subsec:hamiltonian-derivation}

Once projected $\chi$ configurations admit a smooth, stable coarse-grained
description, admissible local relaxation patterns may be summarized by a
constraint formally analogous to a Hamiltonian condition.
The effective local relaxation ordering is bounded by the invariant causal
constant~$c$:
\begin{equation}
  \left(\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}\right)^2
  + \mathcal{V}_{\chi_{\mathrm{eff}}}^2
  = c^2 ,
  \label{eq:effective_hamiltonian_constraint}
\end{equation}
where $\mathcal{V}_{\chi_{\mathrm{eff}}}$ denotes an effective internal
variation functional encoding the resistance of projected configurations to
further relaxation (e.g., solitonic or bound configurations).

Restricting to the monotonic ordering branch yields
\begin{equation}
  \mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}
  = c \sqrt{1
      - \frac{\mathcal{V}_{\chi_{\mathrm{eff}}}^2}{c^2}} ,
  \label{eq:chi_dynamics}
\end{equation}
expressing the universal slowdown of effective relaxation induced by
localized structural constraints.

\subsubsection*{Emergent Gravitational Description}

Projected configurations exhibiting strong resistance to relaxation locally
reduce the admissible effective relaxation rate.
In a spacetime interpretation, this reduction is conventionally described as
gravitational time dilation.
No independent gravitational field is postulated.

In weak-structure regimes, the spatial distribution of relaxation slowdown
admits the simplified relation
\begin{equation}
  \nabla \cdot \left(
    \frac{\nabla \chi_{\mathrm{eff}}}
    {\sqrt{1 - |\nabla \chi_{\mathrm{eff}}|^2 / c^2}}
  \right)
  \simeq \frac{4 \pi G_{\mathrm{eff}}}{c^2} \rho ,
  \label{eq:effective_poisson_relation}
\end{equation}
where $\rho$ denotes the effective density of localized
relaxation-resistant configurations and $G_{\mathrm{eff}}$ is an emergent
coupling parameter.
In the Newtonian limit, this yields an effective Poisson equation for a
potential defined operationally by
\begin{equation}
  \Phi \equiv c^2 \ln \!\left(
    \frac{\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}}{c}
  \right) .
  \label{eq:effective_gravitational_potential}
\end{equation}
This potential is a compact summary of how localized variations in
effective relaxation modulate physical clocks and rulers.
