\subsection{Measurement, Temporal Ordering, and Antiparticle Emergence}
  \label{subsec:measurement-temporal-ordering-and-antiparticle-emergence}

\subsubsection*{Projective Selection and Measurement}

  Let \(\Pi\) denote the non-injective projection from the \(\chi\)-substrate to effective
  observable descriptions.
  A given relational configuration \(\chi_0\) may admit multiple projected descriptions
  \(\{\Pi_\alpha(\chi_0)\}\).

  Measurement corresponds to the stabilization of a single projected description
  \(\Pi_{\alpha^\ast}(\chi_0)\) under interaction with an environment.
  Alternative projected descriptions become inadmissible due to amplification-induced
  constraints.

\subsubsection*{Temporal Ordering from Admissibility}

  Let \(\mathcal{A}\) denote the set of admissible projected configurations.
  Define a partial ordering \(\prec\) such that
  \[
    \Pi_a \prec \Pi_b
  \]
  if \(\Pi_b\) can be obtained from \(\Pi_a\) by admissible relaxation or factorization,
  but not conversely.

  Decay and measurement processes correspond to transitions toward configurations that
  are minimal with respect to \(\prec\).
  This ordering induces an effective arrow of time without introducing a fundamental
  temporal parameter.

\subsubsection*{Structural Origin of Antiparticles}

  Assume that admissible projected configurations carry signed structural invariants
  \(Q \in \mathbb{Z}\) or \(Q \in \mathbb{Z}_2\), associated with orientation, chirality,
  or phase winding.

  For a metastable configuration \(\chi_A\) undergoing admissible factorization,
  invariance requires
  \[
    Q(\chi_A) = \sum_i Q(\chi_i).
  \]

  If \(Q(\chi_A)=0\) but factorization produces nonzero local contributions, admissibility
  requires the appearance of paired configurations with opposite signs,
  \[
    Q(\chi_i) = +q, \quad Q(\chi_j) = -q.
  \]

  These paired configurations are interpreted, at the effective level, as
  particle–antiparticle pairs.
  Their emergence reflects the necessity of preserving signed structural invariants under
  non-injective projection and factorization, rather than the creation of independent
  degrees of freedom.
