\subsection{Neutrino-Mediated Structural Smoothing and Cosmological Inference}
  \label{subsec:neutrino-mediated-structural-smoothing-and-cosmological-inference}

\subsubsection*{Effective Expansion versus Structural Relaxation}

  Let \(a(\tau)\) denote the effective scale factor inferred from projected observables.
  In Cosmochrony, changes in \(a(\tau)\) encode both geometric expansion and cumulative
  structural relaxation.

  We may write, at the effective level,
  \[
    \frac{1}{a}\frac{da}{d\tau}
    =
    H_{\mathrm{geom}}(\tau) + H_{\mathrm{relax}}(\tau),
  \]
  where \(H_{\mathrm{relax}}\) captures the contribution of irreversible structural
  smoothing, prominently mediated by neutrino-like excitations.

\subsubsection*{Early-Time Bias from Neutrino Free Streaming}

  Neutrino-like excitations free-stream over cosmological scales shortly after their
  emission.
  This introduces a non-local smoothing term that suppresses large-scale reconfigurations.

  Early-time observables, such as CMB anisotropies, are sensitive to the integrated
  effect of this smoothing.
  As a result, parameter inference assuming purely geometric expansion may systematically
  underestimate the late-time effective expansion rate.

\subsubsection*{\texorpdfstring{Low-$\ell$}{Low-l} Anomalies as Structural Fossils}

  The lowest multipoles of the CMB correspond to the largest effective scales and
  therefore probe the earliest admissible configurations.

  Suppression or alignment of these modes may reflect anisotropic or directionally
  biased relaxation during the neutrino-dominated smoothing phase.
  Such features are naturally preserved due to the irreversibility of neutrino-mediated
  free streaming.
