% ----------------------------------------------------------------------------
% Section 5.2 --- Gauges as Relaxation Transmittance
% From former ยง10.2
% ----------------------------------------------------------------------------
\subsection{Gauges as Relaxation Transmittance}
\label{sec:gauges-transmittance}

Gauge interactions are reinterpreted as degrees of freedom of the
relaxation flow within the projection fiber~$\Pi$, encoding how locally
admissible projections are coherently related despite non-injectivity.

\begin{itemize}
  \item \textbf{Electromagnetism ($U(1)$):}
    Corresponds to the phase degree of freedom of the relaxation flow
    along the fibers of the Hopf fibration of
    $\Pi \cong S^3$.
    This emergent geometric parameter reflects the freedom to locally
    re-identify projected descriptions without altering global relational
    consistency.
    The fine-structure constant~$\alpha$ characterizes the
    \emph{transmittance} of the relaxation flow through the fiber:
    the stability of admissible projective re-identifications along the
    fiber.

  \item \textbf{Weak Interaction ($SU(2)$):}
    Emerges from the rotational degrees of freedom of the $S^3$
    projection fiber itself.
    These encode non-abelian modes of relaxation transmittance,
    corresponding to shear-like distortions of admissible projections.
    The massive character of the $W^\pm$ and $Z^0$ bosons follows from a
    non-zero spectral gap associated with these modes, interpretable as
    spectral drag or torsion intrinsic to the fiber geometry.

  \item \textbf{Strong Interaction ($SU(3)$):}
    Associated with topological constraints arising from non-trivial
    winding structures of projected configurations.
    The effective $SU(3)$-like symmetry reflects the threefold (triality)
    structure of the minimal self-intersecting stable soliton, identified
    with the trefoil knot ($w=3$).
    Color symmetry emerges as a geometric consequence of topological
    stability classes within the projection fiber.
\end{itemize}
