\subsection{Projective Incompleteness of Standard Physics}
  \label{subsec:projective-incompleteness-standard-physics}

  The comparison with $\Lambda$CDM and the Standard Model can be sharpened
  beyond the phenomenological level.
  Both frameworks describe the structure of the observable space $O$ with
  extraordinary precision, but neither specifies the fundamental space $\Omega$
  nor the projection $\Pi : \Omega \to O$.
  This absence is not a contingent gap to be filled by future work; it is a
  structural incompleteness that each framework signals from within its own
  internal structure, through three independent mechanisms~\cite{Beau2026l}.

  \textbf{(i) Cosmological singularity.}
  The effective metric $g_{\mu\nu}$ ceases to be well-defined for
  $t \lesssim t_{\mathrm{Planck}}$: the Einstein equations break down
  and $g_{\mu\nu}$ is not a valid object.
  A description that fails in a given regime is naturally interpreted as
  arising from a more fundamental structure whose validity extends beyond
  that regime.
  If a projection $\Pi$ relating such a fundamental level $\Omega$ to the
  effective observables is introduced and satisfies the conditions of
  genuine emergence (Section~\ref{subsec:structural-principles-and-projective-regimes}),
  it must be non-injective.
  The singularity is therefore not a physical catastrophe but a structural
  signal: the geometric description has reached the boundary of its
  projectability domain.

  \textbf{(ii) Renormalisation group.}
  The Standard Model is a Wilsonian effective field theory with ultraviolet
  cutoff $\Lambda$.
  The projection $\Pi_\Lambda : \Omega_{\mathrm{UV}} \to O_{\mathrm{IR}}$
  that integrates out modes above $\Lambda$ is non-injective by construction:
  infinitely many UV completions yield the same infrared observables.
  The renormalisation group is therefore not merely a computational tool but
  the explicit acknowledgement of a non-injective projection, with $S_\Pi > 0$
  as a structural necessity.
  In the Cosmochrony framework, the Born--Infeld bound replaces the
  UV regulator with a structurally motivated finite ceiling, rendering
  renormalisation unnecessary as an independent procedure
  (Section~\ref{subsec:projective-origin-of-fluctuations}).

  \textbf{(iii) Quantum measurement.}
  The measurement map $\Pi_{\mathrm{meas}} : \mathcal{H} \to O_{\mathrm{class}}$
  sends distinct quantum states to the same classical outcome.
  The measurement problem is, in the present language, the question of the
  structure of $\Pi_{\mathrm{meas}}$ and why it is non-injective.
  Non-injectivity is not an interpretation added to quantum mechanics: it is
  the formal content of the projection postulate itself.

  These three mechanisms are independent of any hypothesis specific to
  Cosmochrony; they follow from the internal structure of standard physics
  alone.
  Their common conclusion is that $\Lambda$CDM and the Standard Model are
  necessarily projectively incomplete: they describe $O$ without specifying
  the $\Omega$ and $\Pi$ that define their own observables.

  \begin{table}[ht]
    \centering
    \begin{tabular}{lll}
      \hline
      \textbf{Level} & \textbf{Standard physics} & \textbf{Status} \\
      \hline
      Observable space $O$        & Fully described   & present \\
      Fundamental space $\Omega$  & Not specified     & absent  \\
      Projection $\Pi$            & Not specified     & absent  \\
      Fibres $\Pi^{-1}(o)$        & Not specified     & absent  \\
      \hline
    \end{tabular}
    \caption{Projective completeness of standard physics.
    The quantities that $\Lambda$CDM and the Standard Model cannot
    derive from first principles --- the fine-structure constant $\alpha$,
      the electron-to-proton mass ratio $m_e/m_p$, the CKM mixing angles ---
      are precisely the invariants of $\Pi$ that $O$ alone cannot determine:
      they encode properties of the fibres $\Pi^{-1}(o)$ invisible to any
      description confined to the effective level.}
    \label{tab:projective-completeness}
  \end{table}

  The incompleteness is not empirical but structural.
  Standard physics omits the map that defines its own observables.
  Any candidate unification or quantum-gravity programme that remains within
  $O$ --- however sophisticated --- does not complete the description; it
  extends it.
  A theory that genuinely completes standard physics must specify $\Omega$,
  $\Pi$, and the non-injective structure of $\Pi$~\cite{Beau2026l}.
  Cosmochrony provides such a specification.
