\subsection{Flavor Structure and CP Violation}
  \label{subsec:flavor-structure-cp}

  Within the spectral framework established in Chapter~~\ref{sec:spectral-admissibility},
  the interpretation of mass developed above naturally extends to the
  structure of flavor.
  Distinct fermionic generations correspond to spectrally separated
  relaxation modes supported by the same projection fiber~$\Pi$,
  whose geometry was introduced in Section~\ref{sec:geometry-pi}.
  The associated eigenvalues determine the relaxation energy of the
  projected configurations and are therefore directly related to the
  inertial masses introduced in Section~\ref{sec:spectral-stability}.

  \paragraph{Spectral non-degeneracy and the $n=3$ threshold.}

    Let $\{\lambda_i\}$ denote the spectral eigenvalues governing three
    distinct relaxation modes.
    If two eigenvalues become degenerate,
    \[
      \lambda_i = \lambda_j ,
    \]
    the corresponding subspace admits a rotational freedom that allows any
    accumulated phase to be absorbed by a redefinition of the projected
    basis.
    In that case no irreducible CP-odd invariant remains.

    A non-trivial CP-violating sector therefore requires at least three
    spectrally distinct modes.
    The threshold $n \ge 3$ signals that the projective holonomy of the
    flavor sector can no longer be embedded in a two-dimensional manifold.
    An irreducible geometric phase becomes unavoidable once the spectral
    complexity of the fiber exceeds this minimal bound.

  \paragraph{Dual spectral hierarchies.}

    In the Standard Model, CP violation arises from the mismatch between
    the up-type and down-type quark sectors.
    Within Cosmochrony this structure can be interpreted as the
    non-alignment of two sectorial spectral restrictions of the same
    projection fiber.

    Let $\Pi_u$ and $\Pi_d$ denote the effective operators selecting the
    up-type and down-type relaxation modes from the admissible spectral
    subspace of $\Pi \cong S^3$.
    These operators do not define independent geometries;
    they act within the same projection fiber but correspond to distinct
    spectral extractions.

    Because the projected state~$U$ has finite descriptive capacity,
    characterized by the bandwidth constraint~$b_\chi$,
    the two extraction channels cannot in general be simultaneously
    diagonalized.
    This induces a structural non-commutativity,
    \[
      [\Pi_u,\Pi_d] \neq 0 ,
    \]
    reflecting the operational mismatch between the two flavor sectors.

  \paragraph{Cubic commutator invariant.}

    The physically relevant CP-odd quantity can then be expressed through
    the imaginary part of the trace of the cubic commutator,
    \begin{equation}
      \mathcal{J}
      \propto
      \mathrm{Im}\,\mathrm{Tr}\bigl([\Pi_u,\Pi_d]^3\bigr).
    \end{equation}

    The cubic order is minimal for which the invariant vanishes
    identically in a two-generation system and becomes non-trivial only
    when three independent generations are present.
    If the sectorial operators commute,
    \(
    [\Pi_u,\Pi_d]=0 ,
    \)
    then $\mathcal{J}=0$ and CP symmetry is effectively restored.

    When combined with the spectral separations
    $\lambda_{u,i}$ and $\lambda_{d,i}$,
    the invariant scales structurally as
    \[
      \mathcal{J}
      \propto
      \mathrm{Im}\,\mathrm{Tr}([\Pi_u,\Pi_d]^3)
      \prod_{i<j}(\lambda_{u,i}^2-\lambda_{u,j}^2)
      \prod_{k<l}(\lambda_{d,k}^2-\lambda_{d,l}^2).
    \]
    Degeneracy in either sector suppresses the invariant,
    recovering the structural constraint known from the Jarlskog
    formulation of the Standard Model.

  \paragraph{Magnitude of the invariant.}

    The experimentally observed smallness of the Jarlskog invariant,
    $\mathcal{J}\sim3\times10^{-5}$,
    suggests that the flavor sector operates near a
    \emph{quasi-commutative regime} of the projection algebra.

    Let the misalignment between the up-type and down-type operators be
    characterized by a dimensionless parameter $\epsilon\ll1$ such that
    \[
      [\Pi_u,\Pi_d] \sim \epsilon\,K ,
    \]
    where $K$ denotes a normalized operator in the flavor algebra.
    Since the CP-odd quantity is cubic in the commutator,
    \begin{equation}
      \mathcal{J}
      \sim \mathcal{O}(\epsilon^3).
    \end{equation}

    This cubic dependence provides a geometric mechanism for suppressing
    CP violation: even a modest primary misalignment
    (e.g.\ $\epsilon\sim10^{-2}$) can produce an invariant of order
    $\mathcal{J}\sim10^{-6}$, compatible with the observed hierarchy.

  \paragraph{Structural interpretation.}

    Within Cosmochrony, CP violation is therefore not introduced as an
    independent complex phase.
    Rather, it appears as a geometric residue of incompatible spectral
    extractions from a single relational substrate.

    The non-commutativity considered here differs from the
    torsion–relaxation mismatch discussed in
    Section~\ref{sec:noncommutativity-mass}.
    There the commutator acts within a single topological excitation and
    manifests as inertial mass,
    whereas here it arises between distinct sectorial spectral
    restrictions of the same projection fiber and manifests as a
    CP-odd geometric phase.
    Both effects reflect different aspects of the same finite projective
    capacity constraint.
