\subsection{The Geometry of the \texorpdfstring{$\Pi$}{Π} Subspace}
  \label{sec:geometry-pi}

  The relational substrate $\chi$ is not accessed in its full structural
  complexity, but through admissible local projections onto a reduced projection fiber $\Pi \cong S^3$.
  This projection acts as a spectral filter, retaining only those modes of the
  relaxation flow that remain jointly projectable and stable under local re-identification.
  Here, $\Pi$ denotes the projection \emph{fiber}, i.e. the space of admissible internal
  representatives associated with a projected configuration, not the projection map itself.

  The identification $\Pi \cong S^3$ is not imposed \emph{a priori}, but follows from the minimal compact geometry
  (in dimension and connectivity) required to support non-injective yet stable projections.
  It is the minimal compact manifold admitting nontrivial fibrations while preserving
  global stability and boundedness of the projection dynamics.
  Its associated Hopf fibration naturally induces an effective
  $SU(2)\times U(1)$ symmetry structure, which emerges as an invariance of the
  projection process rather than as a fundamental symmetry of the substrate.

  The metric on $\Pi$ is not fixed but dynamically induced by the local density of
  relational connections encoded in the underlying relational graph $G$ used as a numerical and representational
  support, not as a physical discretization.
  The mapping from the global graph Laplacian $\Delta_G$ to the effective projected Laplacian $\Delta_\Pi$ is given by
  \begin{equation}
    \Delta_\Pi = P^\dagger \Delta_G P ,
  \end{equation}
  where $P$ denotes the projection operator onto the admissible spectral subspace.
  The appearance of a smooth three-sphere geometry thus corresponds to a a large-$N$ spectral coarse-graining limit of
  the discrete relational structure, in which coarse-grained spectral properties dominate.
