\subsection{Gauges as Relaxation Transmittance}
  \label{sec:gauges-transmittance}

  Within the Cosmochrony framework, gauge interactions are not fundamental forces
  mediated by independent fields.
  They are reinterpreted as degrees of freedom of the relaxation flow within the
  projection fiber $\Pi$, encoding how locally admissible projections of the
  substrate $\chi$ are coherently related despite the intrinsic non-injectivity of
  the projection.

  \begin{itemize}

    \item \textbf{Electromagnetism ($U(1)$):}
    Electromagnetism corresponds to the phase degree of freedom of the relaxation
    flow along the fibers of the Hopf fibration of $\Pi \cong S^3$.
    This phase reflects the freedom to locally re-identify projected descriptions
    without altering global relational consistency.
    The fine-structure constant $\alpha$ characterizes the effective
    \emph{transmittance} of the relaxation flow through the fiber, quantifying the
    robustness with which phase coherence is maintained across projections.

    \item \textbf{Weak Interaction ($SU(2)$):}
    The weak interaction emerges from the rotational degrees of freedom of the
    $S^3$ projection fiber itself.
    These degrees of freedom encode non-abelian modes of relaxation transmittance,
    corresponding to shear-like distortions of admissible projections.
    The massive character of the $W^\pm$ and $Z^0$ bosons follows from the presence
    of a non-zero spectral gap associated with these modes, which can be interpreted
    as a form of spectral drag or torsion intrinsic to the fiber geometry.

    \item \textbf{Strong Interaction ($SU(3)$):}
    The strong interaction is associated with topological constraints arising from
    non-trivial winding structures of projected configurations.
    In particular, the effective $SU(3)$ symmetry reflects the threefold (triality)
    structure of the minimal self-intersecting stable soliton, identified with the
    trefoil knot ($w=3$).
    Color symmetry thus emerges as a geometric consequence of topological stability
    classes within the projection fiber, rather than as a fundamental internal
    degree of freedom.

  \end{itemize}
