\subsection{Hamiltonian Derivation of the Evolution Equation}
  \label{subsec:hamiltonian-derivation}

  Once projected $\chi$ configurations admit a smooth, stable coarse-grained
  description, admissible local projective ordering patterns may be summarized
  by a constraint formally analogous to a Hamiltonian condition.
  The effective local ordering rate is bounded by the invariant causal
  constant~$c$:
  \begin{equation}
    \left(\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}\right)^2
    + \mathcal{V}_{\chi_{\mathrm{eff}}}^2
    = c^2 ,
    \label{eq:effective_hamiltonian_constraint}
  \end{equation}
  where $\mathcal{V}_{\chi_{\mathrm{eff}}}$ denotes an effective internal
  variation functional encoding the resistance of projectively stable
  configurations to further ordering advancement (e.g., solitonic or
  bound configurations).

  Restricting to the monotonic ordering branch yields
  \begin{equation}
    \mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}
    = c \sqrt{1
    - \frac{\mathcal{V}_{\chi_{\mathrm{eff}}}^2}{c^2}} ,
    \label{eq:chi_dynamics}
  \end{equation}
  expressing the universal slowdown of the effective projective ordering rate
  induced by localized structural constraints.

  \subsubsection*{Emergent Gravitational Description}

    Projectively stable configurations locally reduce the admissible effective
    ordering rate.
    In a spacetime interpretation, this reduction is conventionally described as
    gravitational time dilation.
    No independent gravitational field is postulated.

    In weak-structure regimes, the spatial distribution of ordering slowdown
    admits the simplified relation
    \begin{equation}
      \nabla \cdot \left(
                     \frac{\nabla \chi_{\mathrm{eff}}}
                     {\sqrt{1 - |\nabla \chi_{\mathrm{eff}}|^2 / c^2}}
      \right)
      \simeq \frac{4 \pi G_{\mathrm{eff}}}{c^2} \rho ,
      \label{eq:effective_poisson_relation}
    \end{equation}
    where $\rho$ denotes the effective density of projectively stable
    configurations and $G_{\mathrm{eff}}$ is an emergent coupling parameter.
    In the Newtonian limit, this yields an effective Poisson equation for a
    potential defined operationally by
    \begin{equation}
      \Phi \equiv c^2 \ln \!\left(
                              \frac{\mathcal{D}_{\mathrm{loc}} \chi_{\mathrm{eff}}}{c}
      \right) .
      \label{eq:effective_gravitational_potential}
    \end{equation}
    This potential is a compact summary of how projectively stable configurations
    locally constrain the effective ordering rate, modulating physical clocks
    and rulers.
