\subsection{Spectral Stability and the Emergence of \texorpdfstring{$\hbar_{\mathrm{eff}}$}{ℏeff}}
  \label{sec:hbar_eff_derivation}

  In Cosmochrony, the effective Planck constant \(\hbar_{\mathrm{eff}}\) is not
  introduced as a fundamental quantum postulate.
  Instead, it emerges as a scaling parameter linking spectral stability of
  \(\chi\)-field solitons to effective spacetime observables.

  \subsubsection*{Fundamental scales of the \(\chi\) dynamics}

    The \(\chi\) field is characterized by three independent dynamical scales:
    \begin{itemize}
      \item \(K_0\): maximal relaxation stiffness, with dimensions \([L^{-2}]\),
      \item \(\chi_c\): correlation length at which solitonic configurations stabilize,
      \item \(c\): maximal relaxation speed.
    \end{itemize}

    From these, one may define a natural unit of action associated with the relaxation dynamics,
    \begin{equation}
      \hbar_{\chi}
      \;\equiv\;
      \frac{c^3}{K_0 \chi_c} ,
    \end{equation}
    which has the dimensions of action and is independent of the standard Planck
    constant.
    This quantity is introduced on dimensional grounds as the unique action scale constructible from the fundamental
    relaxation parameters.

    \noindent
    Here and in the following, $K_0$ and $\chi_c$ denote the \emph{bare substrate parameters},
    i.e.\ universal invariants characterizing the rigidity and correlation capacity of the
    $\chi$ field. The scale-dependent values discussed in~\ref{sec:appendix-technical} arise only after
    coarse-graining and do not enter the definition of $\hbar_\chi$.

  \subsubsection*{Spectral origin of effective quantization}

    Quantization in Cosmochrony follows from the discrete spectrum of the stability
    operator \(\Delta_G^{(0)}\).
    For a solitonic excitation with eigenvalue \(\lambda_n\), the characteristic
    frequency of small oscillations scales as
    \begin{equation}
      \nu_n \;\sim\; \frac{c}{\chi_c}\,\sqrt{\lambda_n}\,\mathcal{N}_n^{1/2} .
    \end{equation}

    At the effective spacetime level, identifying the rest energy with the product of this frequency and an effective
    action scale yields
    \begin{equation}
      E_n = \hbar_{\mathrm{eff}}\,\nu_n ,
    \end{equation}
    from which \(\hbar_{\mathrm{eff}}\) emerges as a geometric and spectral quantity,
    not as an independent constant.

  \subsubsection*{Regime-dependent scaling}

    The effective value of \(\hbar_{\mathrm{eff}}\) depends on the scale at which the
    system is probed.
    In regimes where the characteristic spacetime scale
    \(\ell_{\mathrm{spacetime}}\) is comparable to \(\chi_c\),
    \begin{equation}
      \hbar_{\mathrm{eff}} \approx \hbar_{\chi} ,
    \end{equation}
    recovering standard quantum behavior.

    At macroscopic scales \(\ell_{\mathrm{spacetime}} \gg \chi_c\),
    \begin{equation}
      \hbar_{\mathrm{eff}}
      \approx
      \hbar_{\chi}
      \left( \frac{\chi_c}{\ell_{\mathrm{spacetime}}} \right)^2 ,
    \end{equation}
    leading to a strong suppression of quantum effects and the emergence of classical
    behavior.
    This suppression reflects reduced spectral accessibility rather than decoherence or wavefunction collapse.

  \subsubsection*{Consistency with quantum phenomenology}

    In the microscopic regime, where \(\hbar_{\mathrm{eff}} \approx \hbar\), standard
    quantization relations
    \(E = \hbar \nu\) are recovered as effective descriptions.
    This agreement is not postulated but follows from the scaling behavior of
    \(\hbar_{\mathrm{eff}}\) once the projected regime matches laboratory scales.

    \paragraph{Numerical constraints.}

      Reproducing particle-scale quantum behavior requires
    \begin{equation}
      K_0 \chi_c^2 \sim \hbar ,
      \end{equation}
      which constrains the admissible values of the relaxation stiffness and correlation length.
      These constraints are consistent with soliton stability and do not require fine tuning.
      This relation should be read as an order-of-magnitude consistency condition, not as an equality fixing independent
      parameters.
