\subsection{Relation to the Higgs Mechanism: Emergence from
\texorpdfstring{$\chi$}{χ} Dynamics}
  \label{subsec:relation-to-the-higgs-mechanism}

  The Higgs field and its VEV are reinterpreted as effective low-energy
  descriptors of a specific projective regime of~$\chi$.

  \subsubsection*{Structural Transition}
    \label{subsec:emergence-higgs-vev}
    \label{subsec:chi_c-electroweak-scale}
    Below a critical scale~$\chi_c$ (homogeneous regime), only massless
    globally coherent configurations are admissible.
    Above~$\chi_c$ (structured regime), localized relaxation-resistant
    configurations stabilize as massive excitations.
    The electroweak scale is related through
    \begin{equation}
      \langle \phi_H \rangle
      \;\propto\;
      \frac{\hbar_{\mathrm{eff}}\, c}{\chi_c},
    \end{equation}
    naturally recovering the observed scale for $\chi_c \sim 10^{-18}\,\mathrm{m}$.

  \subsubsection*{Mass Generation as Solitonic Stabilization}
    \label{subsec:mass-generation-solitons}
    Fermion masses scale as
    $m_f \propto y_f\,\hbar_{\mathrm{eff}}/\chi_c$;
    gauge boson masses as
    $m_W \propto g\,\hbar_{\mathrm{eff}}/\chi_c$.
    These relations reproduce standard Higgs-generated mass terms at the effective level.

    Within the spectral interpretation developed in Section~\ref{subsec:flavor-structure-cp},
    the Yukawa parameters encode the effective projection of underlying relaxation eigenvalues.
    The observed flavor hierarchies and associated CP structure therefore reflect the same spectral organization,
    rather than introducing independent mass-generating dynamics.

    Within the spectral interpretation developed in Section~\ref{subsec:flavor-structure-cp},
    the Yukawa parameters encode the effective projection of underlying relaxation eigenvalues.
    The observed flavor hierarchies and associated CP structure therefore reflect the same spectral organization,
    rather than introducing independent mass-generating dynamics.

  \subsubsection*{Phenomenological Status}
    \label{subsec:phenomenological-implications}
    \label{subsec:higgs-summary}
    No deviation from established collider results is implied at accessible energies.
    Departures may arise only in extreme regimes.
    Open challenges include deriving the detailed mapping between~$\chi$
    soliton spectra and the full Standard Model mass spectrum.
