\subsection{Saturation selects the Born--Infeld completion}
  \label{subsec:stepB-BI}

  The IR expansion~\eqref{eq:EH-result} is valid only for $\ell_c \gg \ell_\chi$.
  In the ultraviolet regime $\ell_c \sim \ell_\chi$, higher-derivative corrections become
  comparable to the Einstein--Hilbert term, and the expansion breaks down.
  The saturation bound $|\partial_t \chi| \leq c_\chi$ provides the principle that
  selects among the possible UV completions.

  As established in~\cite{Beau2026b}, the minimal conditions for a UV completion
  compatible with bounded flux propagation are:
  \begin{enumerate}[label=(\roman*)]
    \item it reduces to a quadratic (Einstein--Hilbert-like) form in the IR
    ($\ell_c \gg \ell_\chi$),
    \item it enforces a strict upper bound on admissible field invariants at
    $\ell_c \sim \ell_\chi$,
    \item it introduces no additional microscopic scales beyond $c_\chi$ and
    $\hbar_\chi$.
  \end{enumerate}

  Applied to the gravitational sector, a natural Born--Infeld-like completion takes the form:
  \begin{equation}
    S_{\mathrm{coh}}^{\mathrm{full}}[g, \psi]
    = \frac{c_\chi^4}{\hbar_\chi^2}
    \int_{\mathcal{M}} \left(
                         \sqrt{-\det\!\left( g_{\mu\nu}
                                        + \frac{\hbar_\chi}{c_\chi^2} \mathcal{R}_{\mu\nu} \right)}
                         - \sqrt{-g}
    \right) d^4x
    + S_{\mathrm{matter}}[g, \psi],
    \label{eq:BI-gravity}
  \end{equation}
  where $\mathcal{R}_{\mu\nu}$ is a symmetric curvature tensor.
  The precise identification of $\mathcal{R}_{\mu\nu}$ requires the tensorial extension
  of the operator-theoretic analysis of~\cite{Beau2026b}.

  In the IR expansion $\hbar_\chi / c_\chi^2 \ll \ell_c^2$:
  \begin{align}
    \sqrt{-\det\!\left( g_{\mu\nu}
                   + \frac{\hbar_\chi}{c_\chi^2} \mathcal{R}_{\mu\nu} \right)}
    &\approx \sqrt{-g}
    \left(
      1
      + \frac{\hbar_\chi}{2c_\chi^2} \mathcal{R}
      + \mathcal{O}\!\left(\frac{\hbar_\chi^2}{c_\chi^4}
                       \mathcal{R}_{\mu\nu} \mathcal{R}^{\mu\nu}\right)
    \right),
    \label{eq:BI-expand}
  \end{align}
  so that:
  \begin{equation}
    S_{\mathrm{coh}}^{\mathrm{full}}
    \;\xrightarrow{\;\ell_c \gg \ell_\chi\;}\;
    \frac{c_\chi^2}{2\hbar_\chi}
    \int_{\mathcal{M}} \mathcal{R} \sqrt{-g}\, d^4x
    + S_{\mathrm{matter}}
    + \mathcal{O}(\ell_c^{-4}),
    \label{eq:BI-IR}
  \end{equation}
  where $\mathcal{R} \equiv g^{\mu\nu}\mathcal{R}_{\mu\nu}$.
  If $\mathcal{R}_{\mu\nu}=R_{\mu\nu}$, then $\mathcal{R}=R$ and the leading term
  reproduces~\eqref{eq:EH-result}.

  \begin{proposition}[Born--Infeld structure as gravitational UV completion]
    \label{prop:BI-unique}
    Under conditions (i)--(iii) and ontological uniqueness, the UV completion of the
    Einstein--Hilbert term in $S_{\mathrm{coh}}$ belongs to the class of Eddington-inspired
    Born--Infeld actions of the form~\eqref{eq:BI-gravity}.
    This structure is the unique diffeomorphism-invariant extension of the scalar
    Born--Infeld form satisfying (i)--(iii) under the ansatz that the completion depends
    on the metric and its curvature only through a single symmetric rank-2 tensor.
    The tensorial extension of the uniqueness argument from the scalar case to the
    gravitational case remains an open step requiring the full operator analysis.
  \end{proposition}

  \begin{proof}
    The scalar and vector cases are established in~\cite{Beau2026b}.
    The extension to the gravitational sector follows by the same argument under the ansatz
    that the completion depends on $g_{\mu\nu}$ and $\mathcal{R}_{\mu\nu}$ only through their
    determinant combination, which is the minimal diffeomorphism-invariant structure
    satisfying (i)--(iii).
    The identification $\mathcal{R}_{\mu\nu}=R_{\mu\nu}$ is natural but not uniquely forced
    at this level.
    Alternatives such as $\mathcal{R}_{\mu\nu}=R_{\mu\nu}-\frac{1}{4}g_{\mu\nu}R$ are also
    compatible with (i)--(iii) and would modify the IR expansion only at subleading order.
    This residual ambiguity is the primary open point of the gravitational Born--Infeld
    completion within the present framework.
  \end{proof}

  The preceding proposition identifies the Born--Infeld determinantal
  structure as the gravitational UV completion under an explicit ansatz
  on the functional form of the completer.
  We now show that this identification is not an ansatz but a structural
  rigidity: the determinantal form is the \emph{unique} admissible
  spectral tensorial completion compatible with the scalar Born--Infeld
  sector established in~\cite{Beau2026b}.

  \medskip

  \textbf{Setup.}
  Define the BI coupling length
  $\ell^2 = \hbar_\chi / c_\chi^2$
  and the mixed curvature endomorphism
  \begin{equation}\label{eq:mixed-curvature-endo}
    X^{\mu}{}_{\nu}
    \;=\;
    \ell^2\, g^{\mu\alpha}\, \mathcal{R}_{\alpha\nu}\,,
  \end{equation}
  so that the BI action~\eqref{eq:BI-gravity} reads
  $S_{\mathrm{coh}}^{\mathrm{full}}
  = (c_\chi^4/\hbar_\chi^2)\!
  \int \sqrt{-g}\,[\,\Phi(X) - 1\,]\,\mathrm{d}^4 x
  + S_{\mathrm{matter}}$
  with $\Phi(X) = \sqrt{\det(I + X)}$.
  The question is whether $\Phi$ is forced by the admissibility
  conditions, or merely chosen.

  \medskip

  \textbf{Admissibility conditions on the completer.}
  We require that the scalar function $\Phi$ of the
  endomorphism~$X$ satisfy:

  \begin{enumerate}[label=\textbf{(C\arabic*)}]
    \item \label{cond:spectral-cov}
    \textit{Spectral covariance.}\;
    $\Phi(X)$ depends on $X$ only through its eigenvalues:
    $\Phi(U X U^{-1}) = \Phi(X)$ for all
    $U \in \mathrm{O}(d)$.

    \item \label{cond:IR}
    \textit{Einstein infrared limit.}\;
    For weak curvature ($\|X\| \ll 1$):
    $\Phi(I + \varepsilon\, X)
    = 1 + \tfrac{1}{2}\,\varepsilon\,\mathrm{tr}\, X
    + O(\varepsilon^2)$,
    reproducing the Einstein--Hilbert
    density~\eqref{eq:EH-result} at leading order.

    \item \label{cond:saturation}
    \textit{Algebraic Born--Infeld saturation.}\;
    $\Phi(X) \to 0$ as any eigenvalue of $(I + X)$
    approaches zero.
    The completer density vanishes at maximal admissible
    curvature rather than diverging, in accordance with
    the scalar saturation bound
    of~\cite{Beau2026b}.

    \item \label{cond:isotropy}
    \textit{Spectral isotropy.}\;
    $\Phi$ treats all eigenvalue directions symmetrically:
    $\Phi(X) = \varphi(\lambda_1, \ldots, \lambda_d)$
    with $\varphi$ symmetric.

    \item \label{cond:extensivity}
    \textit{Coherence extensivity.}\;
    Define $S_{\mathrm{coh}} = \log \Phi$.
    For admissibly independent sectors $X_1$, $X_2$
    (in the sense of Axiom~A3 of~\cite{Beau2026m}):
    $S_{\mathrm{coh}}(X_1 \oplus X_2)
    = S_{\mathrm{coh}}(X_1)
    + S_{\mathrm{coh}}(X_2)$.
  \end{enumerate}

  Conditions~\ref{cond:spectral-cov}--\ref{cond:isotropy} are
  direct consequences of the covariance, infrared, and
  boundedness properties established in~\cite{Beau2026b} and in
  Section~\ref{subsec:necessity-open}.
  Condition~\ref{cond:extensivity} is the only additional input;
  its derivation from Axiom~A3 is the subject of the following
  hypothesis.

  \begin{hypothesis}[\textbf{H-ext}]\label{hyp:Hext}
    Let $X_1$ and $X_2$ be curvature endomorphisms associated
    with admissibly independent sectors in the sense of
    Axiom~A3: no cross-admissibility constraint couples the
    two sectors after projection locking.
    Then the coherence functional is extensive:
    \begin{equation}\label{eq:H-ext}
      S_{\mathrm{coh}}(X_1 \oplus X_2)
      =
      S_{\mathrm{coh}}(X_1) + S_{\mathrm{coh}}(X_2)\,.
    \end{equation}
  \end{hypothesis}

  \begin{remark}[Structural motivation for \textup{[H-ext]}]
    \label{rem:Hext-motivation}
    Axiom~A3 forbids premature factorisation of the admissible
    fibre prior to projection locking.
    After locking, sectors that carry no mutual admissibility
    constraint contribute independently to the total coherence
    cost.
    The extensivity of $S_{\mathrm{coh}}$ is the post-locking
    counterpart of pre-locking non-factorisability:
    precisely because A3 prevents premature selection,
    the residual cost after locking decomposes additively over
    sectors that have been independently resolved.
    This is structurally analogous to the extensivity of
    thermodynamic entropy for non-interacting subsystems
    and to the additivity of von~Neumann entropy for product
    states.
    An analytical derivation of~\textup{[H-ext]} from~A3
    would promote
    Theorem~\ref{thm:uniqueness-completion} to an unconditional
    result; pending such a derivation,~\textup{[H-ext]}
    occupies a status analogous to~\textup{[H-color]}
    in the spectral admissibility
    sub-programme~\cite{Beau2026a31}.
  \end{remark}

  The uniqueness argument proceeds in three lemmas.
  Lemmas~\ref{lem:1Dreduction}
  and~\ref{lem:spectralcarrier} are unconditional;
  Lemma~\ref{lem:extensivity-consequence} is conditional
  on~[H-ext] and carries the only conditionality in the
  chain.

  \begin{lemma}[Coherence extensivity forces spectral
  decomposition]
    \label{lem:extensivity-consequence}
    Assume~\textup{[H-ext]}.
    Then conditions~\ref{cond:spectral-cov},
    \ref{cond:isotropy},
    and~\ref{cond:extensivity} together force
    \begin{equation}\label{eq:spectral-decomposition}
      S_{\mathrm{coh}}(X)
      = \sum_{i=1}^{d} h(\lambda_i)\,,
    \end{equation}
    for a single smooth function
    $h : (-1, +\infty) \to \mathbb{R}$, where
    $\lambda_1, \ldots, \lambda_d$ are the eigenvalues
    of~$X$.
  \end{lemma}

  \begin{proof}
    The extensivity of $S_{\mathrm{coh}}$
    (Hypothesis~[H-ext]) implies
    $\Phi(X_1 \oplus X_2)
    = \Phi(X_1)\,\Phi(X_2)$
    by exponentiation.
    Write $X$ in diagonal form
    $X = \mathrm{diag}(\lambda_1, \ldots, \lambda_d)$
    relative to the $g$-orthonormal eigenbasis.
    Apply the multiplicativity of $\Phi$ iteratively to
    the rank-1 decomposition
    $X = \lambda_1 P_1 \oplus \cdots \oplus \lambda_d P_d$
    (where $P_i$ are rank-1 projectors onto the
    eigenspaces):
    $\Phi(X) = \prod_{i} \Phi(\lambda_i P_i)$.
    Taking logarithms gives~\eqref{eq:spectral-decomposition}
    with $h(\lambda) = \log \Phi(\lambda P)$ for any unit
    projector~$P$.
    Condition~\ref{cond:isotropy} ensures that $h$ does not
    depend on the choice of eigendirection.
  \end{proof}

  \begin{lemma}[One-dimensional reduction to scalar BI]
    \label{lem:1Dreduction}
    Under
    conditions~\ref{cond:IR} and~\ref{cond:saturation},
    the sector function $h$
    in~\eqref{eq:spectral-decomposition} is uniquely
    determined:
    \begin{equation}\label{eq:h-unique}
      h(\lambda)
      = \tfrac{1}{2}\,\log(1 + \lambda)\,.
    \end{equation}
  \end{lemma}

  \begin{proof}
    \textit{Infrared constraint.}\;
    Condition~\ref{cond:IR} requires, at first order:
    $\Phi(I + \varepsilon X)
    = \exp\!\bigl(\sum_i h(\varepsilon \lambda_i)\bigr)
    \approx 1 + \varepsilon\, h'(0)\,\mathrm{tr}\, X$.
    Matching with
    $1 + \tfrac{1}{2}\,\varepsilon\,\mathrm{tr}\, X$ gives
    $h'(0) = 1/2$.
    Since $\Phi(0) = 1$ at flat space, $h(0) = 0$.

    \textit{Saturation constraint.}\;
    Condition~\ref{cond:saturation} requires
    $\Phi(X) \to 0$ as $\lambda \to -1^+$,
    hence $h(\lambda) \to -\infty$ as
    $\lambda \to -1^+$.

    \textit{Reduction to the scalar sector.}\;
    In the scalar Born--Infeld theory derived
    in~\cite{Beau2026b}, the one-dimensional completer is
    uniquely fixed by the saturation bound:
    the action density takes the form
    $\sqrt{1 + \lambda} - 1$ for a single mode of
    normalised field gradient~$\lambda$.
    The associated coherence density is
    $\log \sqrt{1 + \lambda}
    = \tfrac{1}{2}\log(1 + \lambda)$,
    satisfying $h(0) = 0$, $h'(0) = 1/2$,
    and $h \to -\infty$ as $\lambda \to -1^+$.

    \textit{Uniqueness.}\;
    Any alternative $\tilde{h}$ satisfying the above can be
    written as
    $\tilde{h}(\lambda)
    = \tfrac{1}{2}\log(1 + \lambda) + r(\lambda)$
    with $r(0) = 0$, $r'(0) = 0$, and $r$ bounded below
    near $\lambda = -1$.
    In the one-dimensional sector, $\exp(\tilde{h})$ must
    reproduce the scalar BI density $\sqrt{1 + \lambda}$,
    whence $\exp(r(\lambda)) = 1$ identically and
    $r \equiv 0$.
  \end{proof}

  \begin{lemma}[Spectral carrier identification]
    \label{lem:spectralcarrier}
    Among local covariant two-derivative symmetric tensors
    on $(M, g)$, the mixed Ricci endomorphism
    $X^{\mu}{}_{\nu}
    = \ell^2\, g^{\mu\alpha} R_{\alpha\nu}$
    is the unique admissible spectral carrier for~$\Phi$
    with non-degenerate eigenvalue spectrum.
  \end{lemma}

  \begin{proof}
    \textit{Classification.}\;
    At two-derivative order, the only symmetric $(0,2)$
    tensors constructed covariantly from the metric and
    the Riemann tensor are $R_{\mu\nu}$ and
    $R\, g_{\mu\nu}$, together with linear combinations.
    Condition~\ref{cond:spectral-cov} requires the
    argument of~$\Phi$ to be an endomorphism
    $X : T_p M \to T_p M$, obtained by raising one
    index with the metric.

    \textit{Exclusion of
      $f(R)\,\delta^{\mu}{}_{\nu}$.}\;
    The endomorphism
    $X = f(R)\,\delta^{\mu}{}_{\nu}$ has a single
    repeated eigenvalue.
    It carries no directional spectral information:
    two manifolds with equal scalar curvature but
    distinct Ricci eigenvalue distributions would
    produce identical~$\Phi$, violating the spectral
    resolution required by admissibility.
    In particular, the resulting completer
    $\Phi(X) = \exp\!\bigl(d \cdot h(f(R))\bigr)$ is a
    scalar function of~$R$ alone, degenerate along
    all eigenvalue directions.

    \textit{Selection.}\;
    The remaining candidate is the mixed Ricci tensor
    $g^{\mu\alpha} R_{\alpha\nu}$, which carries the
    full directional information of the Ricci curvature
    through its $d$ independent eigenvalues.
    It is the unique minimal two-derivative endomorphism
    with non-trivial spectral content.

    \textit{Trace modifications.}\;
    The alternative
    $\mathcal{R}_{\mu\nu}
    = R_{\mu\nu} - \alpha\, g_{\mu\nu} R$ (for any
    constant~$\alpha$) yields an endomorphism with
    eigenvalues shifted by $-\alpha R$ relative to the
    Ricci eigenvalues.
    Such a shift modifies $\Phi$ only at subleading IR
    order: the leading term
    $\tfrac{1}{2}\,\mathrm{tr}\, X
    = \tfrac{1}{2}\,(1 - d\alpha)\,\ell^2 R$ is
    absorbed into the normalisation of
    $\ell^2$ provided $\alpha \neq 1/d$.
    The identification
    $\mathcal{R}_{\mu\nu} = R_{\mu\nu}$
    (i.e.\ $\alpha = 0$) is the minimal choice, involving
    no trace subtraction; the residual ambiguity
    in~$\alpha$ affects only higher-order corrections
    and is resolved by matching to the heat-kernel
    coefficient $a_2$ of the Gravity
    paper~\cite{Beau2026h}.
  \end{proof}

  \begin{theorem}[Uniqueness of the admissible tensorial
  completion]
    \label{thm:uniqueness-completion}
    Let $\mathcal{C}(g, \mathcal{R})$ be a local covariant
    scalar density of the form
    $\sqrt{-g}\,[\Phi(X) - 1]$, satisfying
    conditions~\ref{cond:spectral-cov}--\ref{cond:isotropy}
    and Hypothesis~\textup{[H-ext]}.
    Then, up to the trace ambiguity resolved by the
    Gravity paper~\cite{Beau2026h}:
    \begin{equation}\label{eq:unique-completer}
      \boxed{\;
      \mathcal{C}(g, R)
      \;=\;
      \sqrt{-g}\;
      \Bigl[\,
      \sqrt{\det\!\bigl(
      \delta^{\mu}{}_{\nu}
      + \ell^2\, g^{\mu\alpha} R_{\alpha\nu}
      \bigr)}
      \;-\; 1
      \,\Bigr]\,.
      \;}
    \end{equation}
  \end{theorem}

  \begin{proof}
    By Lemma~\ref{lem:spectralcarrier}, the spectral
    carrier is
    $X = \ell^2\, g^{-1} R$ with eigenvalues
    $\lambda_i$.
    By Lemma~\ref{lem:extensivity-consequence}
    (conditional on~[H-ext]),
    $\Phi(X) = \exp\!\bigl(\sum_i h(\lambda_i)\bigr)$.
    By Lemma~\ref{lem:1Dreduction},
    $h(\lambda) = \tfrac{1}{2}\log(1 + \lambda)$.
    Therefore:
    \begin{equation*}
      \Phi(X)
      = \exp\!\Bigl(\,
      \tfrac{1}{2}\sum_i \log(1 + \lambda_i)
      \Bigr)
      = \sqrt{\,\prod_i (1 + \lambda_i)}
      = \sqrt{\det(I + X)}\,.
    \end{equation*}
    Substituting into the completer form
    gives~\eqref{eq:unique-completer}.
  \end{proof}

  \begin{remark}[Eddington--Born--Infeld form]
    \label{rem:EBI-form}
    Using $\det(g \cdot A) = \det g \cdot \det A$, the
    action~\eqref{eq:BI-gravity} with
    $\mathcal{R}_{\mu\nu} = R_{\mu\nu}$ becomes
    \begin{equation*}
      S_{\mathrm{coh}}^{\mathrm{full}}[g]
      =
      \frac{c_\chi^4}{\hbar_\chi^2}
      \int_{\mathcal{M}}
      \Bigl[\,
      \sqrt{-\det\!\bigl(
      g_{\mu\nu} + \ell^2\, R_{\mu\nu}
      \bigr)}
      - \sqrt{-g}
      \,\Bigr]\,\mathrm{d}^4 x\,,
    \end{equation*}
    which coincides in form with Eddington-inspired
    Born--Infeld (EiBI) gravitational actions studied
    independently in the modified gravity literature.
    The present derivation differs in that the EiBI form
    is not postulated but derived as a structural rigidity
    from admissibility conditions, and the length
    $\ell = \sqrt{\hbar_\chi / c_\chi^2}$ is not a free
    parameter but is determined by the same spectral
    cutoff that fixes the induced Newton constant.
  \end{remark}

  \begin{remark}[Infrared and ultraviolet spectral
  hierarchy]
    \label{rem:IR-UV-hierarchy}
    The framework involves two spectrally defined objects
    that must be carefully distinguished.
    At the \emph{global} level, the projective entropy
    functional
    $S_\Pi[g] = \tfrac{1}{2}\log\det' A_g$
    encodes the full eigenvalue spectrum of the effective
    Laplacian~$A_g$ and produces the infrared Einstein
    response (Section~\ref{subsec:necessity-open}).
    At the \emph{local} level, the coherence density
    $S_{\mathrm{coh}}(X) = \tfrac{1}{2}\log\det(I + X)$
    encodes the eigenvalue spectrum of the mixed Ricci
    endomorphism and governs the ultraviolet saturation
    regime where the Einstein--Hilbert expansion breaks
    down.
    Both share the algebraic form
    $\tfrac{1}{2}\log\det(\cdot)$ because both originate
    from the extensivity of admissible coherence:
    the former over eigenvalue modes of~$A_g$, the latter
    over eigenvalue directions of~$g^{-1}R$.
    Admissible spectral extensivity is therefore the
    common principle underlying both the infrared
    gravitational response and the ultraviolet
    Born--Infeld completion.
  \end{remark}

  \begin{remark}[Epistemic status]
    \label{rem:epistemic-BI}
    Lemma~\ref{lem:1Dreduction} is unconditional: it
    follows from the scalar Born--Infeld uniqueness
    derived in~\cite{Beau2026b}.
    Lemma~\ref{lem:spectralcarrier} is unconditional:
    it is a classification result in Riemannian geometry.
    Lemma~\ref{lem:extensivity-consequence} is conditional
    on Hypothesis~[H-ext], whose derivation from
    Axiom~A3 is structurally motivated
    (Remark~\ref{rem:Hext-motivation}) but not yet
    analytically closed.
    Theorem~\ref{thm:uniqueness-completion} is therefore
    conditional on~[H-ext] alone.
    The conditionality is sharply localised: it resides
    entirely in the passage from non-factorisability
    (Axiom~A3) to extensivity of
    $S_{\mathrm{coh}}$.
    Proposition~\ref{prop:BI-unique} is now subsumed as
    a consequence of
    Theorem~\ref{thm:uniqueness-completion}: the ansatz
    on the determinantal form is no longer an independent
    assumption but a derived structural rigidity.
  \end{remark}
