\subsection{Strong Gravitational Lensing}
\label{subsec:strong-gravitational-lensing}

The effective lensing potential decomposes as
$\Phi_{\mathrm{eff}} = \Phi_{\mathrm{bar}} + \Phi_{\chi}$,
with convergence
\begin{equation}
  \kappa(\boldsymbol{\theta})
  = \frac{D_l D_{ls}}{c^2 D_s}
    \int \nabla_\perp^2
    \Phi_{\mathrm{eff}}\!
      \left(D_l\boldsymbol{\theta},z\right)\,dz ,
\end{equation}
yielding $\kappa = \kappa_{\mathrm{bar}} + \kappa_{\chi}$.
The emergent contribution is parametrized as
\begin{equation}
  \Phi_{\chi}(\mathbf{x})
  = \frac{c^2}{2}\,
    \ln\!\left(
      \frac{K(\bar\chi(\mathbf{x}))}{K_\infty}
    \right).
\end{equation}
Testable signatures include partial decorrelation between baryonic mass
and lensing strength, enhanced sensitivity to cluster morphology, strong
lensing without dark matter substructure, and redshift dependence tracing
relaxation rather than mass accretion history.

\subsubsection*{Early Massive Structures in the JWST Era}
\label{subsec:jwst-early-structures}

Massive galaxies and strong lensing at $z \gtrsim 8$ are naturally
accommodated: effective mass reflects localized resistance to relaxation
associated with spectrally robust projected configurations, rather than
requiring rapid hierarchical assembly.

\subsubsection*{``Impossibly Early'' Galaxies}
\label{subsec:impossibly-early-galaxies}

Their large effective masses encode strong resistance to relaxation rather
than cumulative accretion.
Cosmochrony predicts that such galaxies should exhibit relatively stable
effective masses over extended redshift intervals, contrasting with the
rapid growth required in hierarchical scenarios.

\subsubsection*{Qualitative Predictions}
\label{subsec:qualitative-prediction-early-galaxies}
\label{subsec:qualitative-prediction-early-lensing}

Massive galaxies at $z \gtrsim 8$ should persist with only moderate mass
evolution.
Early strong-lensing configurations should display weak redshift evolution
of their effective convergence, with an early onset followed by moderate
evolution.

\subsubsection*{Strong Gravitational Lensing in Abell-1689}
\label{subsec:a1689-strong-lensing}

The emergent convergence is isolated as
\begin{equation}
  \kappa_{\chi}(\boldsymbol{\theta})
  = \kappa_{\mathrm{eff}}(\boldsymbol{\theta})
    - \kappa_{\mathrm{bar}}(\boldsymbol{\theta}),
\end{equation}
with the emergent lensing potential obtained from
$\nabla_{\boldsymbol{\theta}}^2 \psi_{\chi} =
  2\kappa_{\chi}$.
In Cosmochrony, $\kappa_{\chi}$ is emergent geometric focusing from
collective $\chi$-substrate constraints, not additional dark matter.

  \subsubsection*{Quantitative Scaling and Falsifiability}

    Beyond qualitative decorrelation effects, the emergent contribution
    $\kappa_{\chi}$ admits an order-of-magnitude scaling in saturated
    cluster cores.

    We parameterize the relative excess convergence as
    \begin{equation}
      \frac{\kappa_{\chi}}{\kappa_{\mathrm{bar}}}
      \sim \gamma
      \left(
        \frac{\Sigma_{\mathrm{bar}}}
        {\Sigma_*}
      \right),
    \end{equation}
    where $\Sigma_{\mathrm{bar}}$ is the baryonic surface density,
    $\Sigma_*$ denotes a characteristic structural relaxation scale of
    the $\chi$ substrate, and $\gamma$ is a dimensionless response
    coefficient encoding substrate efficiency.

    For massive clusters with
    $\Sigma_{\mathrm{bar}} \sim \Sigma_*$,
    this yields
    \begin{equation}
      \frac{\kappa_{\chi}}{\kappa_{\mathrm{bar}}}
      \sim \gamma ,
    \end{equation}
    with a natural expectation
    $\gamma \sim 10^{-2} \text{--} 10^{-1}$,
    corresponding to percent-to-ten-percent level decorrelation
    between baryonic mass distribution and total lensing convergence.

    This deviation should manifest as a systematic mismatch between
    high-resolution baryonic mass reconstructions and gravitational
    lensing maps, particularly in morphologically complex cluster cores
    such as Abell-1689.

    \textbf{Falsifiability condition.}
    If future lensing reconstructions constrain
    $\kappa_{\chi}/\kappa_{\mathrm{bar}} < 10^{-2}$
    in saturated cluster environments, the present scaling law
    would be excluded at the predicted amplitude level.
