% ============================================================================
% Chapter 1 --- Introduction and Motivation
% Merges former sections 1 and 2 of Cosmochrony v1.13beta1
% ============================================================================
\clearpage

\section{Introduction and Motivation}
\label{sec:introduction}

Modern fundamental physics is built upon two highly successful yet conceptually
distinct frameworks: quantum mechanics and general
relativity~\cite{Dirac1930,Einstein1915}.
Quantum theory accurately describes microscopic phenomena, while general
relativity provides a geometric account of gravitation and spacetime dynamics
at macroscopic and cosmological scales.
Despite their empirical success, these theories rely on incompatible
foundational assumptions and resist unification within a single coherent
conceptual framework~\cite{MisnerThorneWheeler1973,Weinberg1972,rovelli2004quantum}.

In this work, we introduce \emph{Cosmochrony}\footnote{%
  From \textit{$\kappa\acute{o}\sigma\mu o\varsigma$} and
  \textit{$\chi\rho\acute{o}\nu o\varsigma$}, denoting a framework in which
  cosmic structure and temporal ordering emerge from a common pre-geometric
  substrate.},
a deliberately minimalist framework whose guiding hypothesis is that spacetime
geometry, gravitation, and quantum phenomena emerge from the dynamics of a
single continuous underlying entity, denoted~$\chi$, whose effective
descriptions arise through a constrained projection process.
This projection is generically non-injective, allowing distinct underlying
$\chi$-configurations to correspond to identical effective observables and,
conversely, allowing a single underlying configuration to admit multiple
correlated effective realizations.
A detailed and formal treatment of this projection asymmetry is given in
Section~\ref{subsec:projection-reality-and-ontological-asymmetry}.

The substrate~$\chi$ is not defined on a pre-existing spacetime manifold, nor
is it interpreted as a conventional physical field propagating within
spacetime.
Instead, spacetime notions themselves arise as effective and relational
descriptions, applicable only once suitable stability and projection conditions
are satisfied.
The precise ontological status of~$\chi$ and the minimal assumptions governing
its dynamics are introduced systematically in
Section~\ref{sec:definition-and-fundamental-properties-of-the-chi-field}.

The fundamental dynamical postulate of Cosmochrony is that $\chi$ undergoes an
irreversible relaxation process, locally bounded by an invariant structural
propagation speed.
The effective projection of this bound defines the observed causal limit~$c$
and induces an intrinsic ordering of physical processes, identified with
physical time.
Spatial relations emerge relationally from differences, gradients, and
correlations of~$\chi$ once a stable geometric regime is reached.
Within this perspective, spacetime expansion, gravitation, particle-like
excitations, radiation processes, and quantum correlations are emergent
phenomena associated with specific configurations or interactions of the
underlying substrate.
In particular, discreteness, inertial mass, and quantum indeterminacy arise
from structural constraints on projection and relaxation.

\begin{conventionbox}{Convention on spacetime language}
  Throughout this article, spacetime coordinates, metric quantities,
  variational principles, and differential geometry are employed strictly as
  effective descriptive tools, valid only in regimes where $\chi$-configurations
  admit a stable geometric interpretation.
  They are not treated as fundamental postulates of the theory.
  Technical reconstructions and mathematical details are confined to the
  appropriate effective regimes and collected in the appendices.
  This point will not be reiterated in subsequent sections.
\end{conventionbox}

Cosmochrony does not aim to replace the Standard Model or general relativity
in their empirically validated domains, nor does it claim to provide a final
unification of quantum theory and gravitation.
It offers an exploratory and internally coherent framework designed to clarify
the physical origin of time, geometry, gravitation, and quantum correlations
within a single relational dynamics.
The unifying thread is the idea that apparent multiplicity, indeterminacy, and
nonlocality reflect structural features of projection.

\input{part1/ch01/sec-unification}
\input{part1/ch01/sec-minimalism}
\input{part1/ch01/sec-time}
\input{part1/ch01/sec-context}
\input{part1/ch01/sec-scope}
\input{part1/ch01/sec-plan}

For convenience, a glossary summarizing the main quantities and operators used
throughout the article is provided in Appendix~\ref{appendix:glossary}.
