\subsection{Emergent Curvature}
  \label{subsec:emergent-curvature}

  Spatial variations in admissible relaxation ordering lead to non-uniform
  correlation patterns within projected descriptions.
  When a smooth geometric parametrization is applicable, these
  non-uniformities are summarized by gradients of an emergent metric
  structure, reproducing the phenomenology traditionally attributed to
  curved spacetime: gravitational time dilation, geodesic deviation, and
  lensing effects.

  \paragraph{Einstein's Equations as a Structural Equilibrium Principle.}
    \label{subsec:einstein-equilibrium-principle}
    When projected configurations admit a smooth, locally injective geometric
    description, their collective structural constraints can be summarized by
    an effective metric~$g_{\mu\nu}$.
    Einstein's equations emerge \emph{necessarily} as the unique
    generally covariant, low-derivative consistency condition relating
    curvature to the effective distribution of relaxation-resistant
    configurations in the macroscopic projected regime.
    The Einstein tensor encodes a geometric identity constraining admissible
    macroscopic descriptions; the stress--energy tensor summarizes how
    localized configurations resist relaxation ordering.
    Einstein's equations therefore express a balance condition between
    geometry and physical structure.

    A direct infrared derivation of the Einstein response from the spectral
    entropy functional of the relational operator is developed in the
    companion work~\cite{Beau2026h}.
    The present section focuses on the effective phenomenological
    interpretation within the Cosmochrony framework.

    This interpretation explains the extraordinary universality of general
    relativity: as long as a smooth spacetime description exists and
    relaxation ordering is monotonic, bounded, and weakly inhomogeneous, the
    same geometric relations must hold, independently of the microscopic
    nature of the underlying substrate.
    General relativity thus plays a role analogous to thermodynamics---exact
    within its domain, silent outside it.

    When projection ceases to be locally injective, Einstein's equations do
    not break down; they simply no longer apply, because the concept of
    spacetime has not yet emerged.
